---
title: "Condition model"
author: "Max Lindmark, Sean C. Andersson, Mayya Gogina and Michele Casini"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align ='center'
)
```

# Fit condition model with environmental and biological predictors
Fit main model (see exploratory scripts and model comparison), visualize results.

## Load packages

```{r packages, message=FALSE, warning=TRUE}
library(tidyverse)
library(tidylog)
library(viridis)
library(marmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(RColorBrewer)
library(gganimate)
library(gifski)
library(latex2exp)
library(patchwork)
library(png)
library(RCurl)
library(wesanderson)
library(qwraps2) 
library(ggcorrplot)
library(ggridges)
library(visreg)
# remotes::install_github("pbs-assess/sdmTMB")
library(sdmTMB)

# To load entire cache in interactive r session, do: 
# qwraps2::lazyload_cache_dir(path = "R/supp_analysis/spatial_sd_sub_div_cache/html")
```

## Code for map plots

```{r read coastline data, message=FALSE, warning=FALSE, fig.show='hide'}
# Define plotting theme for main plot
theme_plot <- function(base_size = 11, base_family = "") {
  theme_light(base_size = base_size, base_family = "") +
    theme(
      axis.text = element_text(size = 9),
      legend.text = element_text(size = 9),
      legend.title = element_text(size = 9),
      legend.position = "bottom",
      legend.key.height = unit(0.2, "cm"),
      legend.margin = margin(0, 0, 0, 0),
      legend.box.margin = margin(-5, -5, -5, -5),
      strip.text = element_text(size = 9, colour = 'gray10', margin = margin(b = 1, t = 1)),
      strip.background = element_rect(fill = "grey95")
      )
}

```

## Read data

```{r read and process data, message=FALSE, warning=FALSE, fig.show='hide'}
# Read the split files and join them
d1 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod-condition/master/data/for_analysis/mdat_cond_(1_2).csv")
d2 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod-condition/master/data/for_analysis/mdat_cond_(2_2).csv")

d <- bind_rows(d1, d2)

unique(is.na(d$density_cod))
unique(is.na(d$density_cod_rec))

# Calculate standardized variables
d <- d %>% 
  mutate(ln_length_cm = log(length_cm),
         ln_weight_g = log(weight_g),
         year_ct = year - mean(year),
         biomass_her_sc = biomass_her,
         biomass_her_sd_sc = biomass_her_sd,
         biomass_spr_sc = biomass_spr,
         biomass_spr_sd_sc = biomass_spr_sd,
         density_cod_sc = density_cod,
         density_cod_rec_sc = density_cod_rec,
         density_fle_sc = density_fle,
         density_fle_rec_sc = density_fle_rec,
         density_saduria_sc = density_saduria,
         density_saduria_rec_sc = density_saduria_rec,
         depth_sc = depth,
         depth_rec_sc = depth_rec,
         oxy_sc = oxy,
         oxy_rec_sc = oxy_rec,
         temp_sc = temp,
         temp_rec_sc = temp_rec,
         year_f = as.factor(year))  %>%
  mutate_at(c("biomass_her_sc", "biomass_her_sd_sc", "biomass_spr_sc", "biomass_spr_sd_sc",
              "density_cod_sc", "density_cod_rec_sc", 
              "density_fle_sc", "density_fle_rec_sc", 
              "density_saduria_sc", "density_saduria_rec_sc", 
              "depth_sc", "depth_rec_sc",
              "oxy_sc", "oxy_rec_sc",
              "temp_sc", "temp_rec_sc"
              ),
            ~(scale(.) %>% as.vector)) %>% 
  mutate(year = as.integer(year))

unique(is.na(d))

unique(is.na(d$density_cod_rec))
unique(is.na(d$density_cod))

# Drop NA covariates for the correlation plot
d <- d %>% drop_na(biomass_her_sc, biomass_her_sd_sc, biomass_spr_sc, biomass_spr_sd_sc,
                   density_cod_sc, density_cod_rec_sc, density_fle_sc, density_fle_rec_sc, 
                   density_saduria_sc, density_saduria_rec_sc, depth_sc, depth_rec_sc,
                   oxy_sc, oxy_rec_sc, temp_sc, temp_rec_sc)
```

## Make spde mesh

```{r make mesh, results='hide', message=FALSE}
spde <- make_mesh(d, xy_cols = c("X", "Y"),
                  n_knots = 100,  
                  type = "kmeans", seed = 42)
```

## Fit models
### Non spatial model to get the average length weight relationship. The main model's response variable is the ratio of the observed vs this predicted average weight.

```{r first fit non_spatial model without covariates, cache=TRUE, fig.show='hide'}
ptm <- proc.time()
mnull <- sdmTMB(
  formula = ln_weight_g ~ ln_length_cm,
  data = d,
  time = NULL,
  mesh = spde,
  family = student(link = "identity", df = 5),
  spatiotemporal = "off",
  spatial = "off",
  silent = TRUE,
  reml = FALSE
  )
proc.time() - ptm

# Extract average a and b across the total dataset
a <- as.numeric(tidy(mnull)[1, "estimate"])
b <- as.numeric(tidy(mnull)[2, "estimate"])
a
b

d$weight_g_avg_pred <- exp(a)*d$length_cm^b

plot(d$weight_g ~ d$length_cm)
plot(d$weight_g_avg_pred ~ d$length_cm)
plot(d$weight_g_avg_pred ~ d$weight_g); abline(a = 0, b = 1, col = "red")

d$le_cren <- d$weight_g / d$weight_g_avg_pred
d$log_le_cren <- log(d$le_cren)
hist(d$le_cren)
```

### Fit main model

```{r spatiotemporal le cren model, cache=TRUE}
ptm <- proc.time()

d$sub_div_f <- as.factor(d$sub_div)

mfull <- sdmTMB(
  formula = log_le_cren ~ -1 + sub_div_f + year_f + biomass_her_sc + biomass_her_sd_sc +
    biomass_spr_sc*sub_div_f + biomass_spr_sd_sc +
    density_cod_sc*sub_div_f + density_cod_rec_sc + 
    density_fle_sc + density_fle_rec_sc + 
    density_saduria_sc + density_saduria_rec_sc*sub_div_f + 
    depth_sc + depth_rec_sc + 
    oxy_sc + oxy_rec_sc*sub_div_f + 
    temp_sc + temp_rec_sc,
    data = d,
  time = "year",
  mesh = spde, 
  family = student(link = "identity", df = 5),
  spatiotemporal = "iid",
  spatial = "on",
  silent = TRUE,
  reml = FALSE,
  control = sdmTMBcontrol(newton_loops = 1)
  )
system("say Just finished!")
proc.time() - ptm

sanity(mfull)

tidy(mfull, conf.int = TRUE) %>% arrange(desc(abs(estimate)))
tidy(mfull, effects = "ran_par", conf.int = TRUE) 
```

Rename and load the main model

```{r}
minter <- mfull

mfull <- readRDS("output/mfull.rds")

print(minter)
print(mfull)

AIC(mfull, minter)
```

Plot conditional effects with visreg, add in the main effects from the full model. Very large differences by sd?

```{r plot coefficients}
#summary(lm(log_le_cren ~ year_f + density_saduria_rec_sc, data = filter(d, sub_div == 24)))

tidy(minter, effects = "ran_par", conf.int = TRUE) %>%
  filter(term %in% c("sigma_O", "sigma_E"))

tidy(mfull, effects = "ran_par", conf.int = TRUE) %>%
  filter(term %in% c("sigma_O", "sigma_E"))


tidy(mfull, conf.int = TRUE) %>% as.data.frame()
tidy(minter, conf.int = TRUE) %>% as.data.frame()

tidy(mfull, conf.int = TRUE) %>% filter(grepl('saduria', term))
tidy(minter, conf.int = TRUE) %>% filter(grepl('saduria', term))


# Extract random and fixed coefficients from the full model
minter_est1 <- tidy(minter, effects = "fixed", conf.int = TRUE) %>% 
  filter(grepl(':', term)) %>%
  mutate() %>% 
  mutate(group = ifelse(grepl("saduria", term), "Saduria (rec)", NA),
         group = ifelse(grepl("cod", term), "Cod (rec)", group),
         group = ifelse(grepl("spr", term), "Sprat (rec)", group),
         group = ifelse(grepl("oxy", term), "Oxygen (rec)", group))

minter_est2 <- tidy(minter, effects = "fixed", conf.int = TRUE) %>% 
  filter(term %in% c("biomass_spr_sc", "density_cod_sc", "density_saduria_rec_sc", "oxy_rec_sc")) %>%
  mutate(term = ifelse(term == "biomass_spr_sc", "sub_div_f24:biomass_spr_sc", term),
         term = ifelse(term == "density_cod_sc", "sub_div_f24:density_cod_sc", term),
         term = ifelse(term == "density_saduria_rec_sc", "sub_div_f24:density_saduria_rec_sc", term),
         term = ifelse(term == "oxy_rec_sc", "sub_div_f24:oxy_rec_sc", term)) %>%
  mutate(group = ifelse(grepl("saduria", term), "Saduria (rec)", NA),
         group = ifelse(grepl("cod", term), "Cod (rec)", group),
         group = ifelse(grepl("spr", term), "Sprat (rec)", group),
         group = ifelse(grepl("oxy", term), "Oxygen (rec)", group))

minter_est2

minter_est <- bind_rows(minter_est1, minter_est2)

minter_est <- minter_est %>%
  mutate(est2 = estimate,
         est2 = ifelse(term %in% c("sub_div_f25:biomass_spr_sc",
                                   "sub_div_f26:biomass_spr_sc",
                                   "sub_div_f27:biomass_spr_sc",
                                   "sub_div_f28:biomass_spr_sc"),
                       filter(minter_est2, term == "sub_div_f24:biomass_spr_sc")$estimate + estimate,
                       est2),

         est2 = ifelse(term %in% c("sub_div_f25:oxy_rec_sc",
                                   "sub_div_f26:oxy_rec_sc",
                                   "sub_div_f27:oxy_rec_sc",
                                   "sub_div_f28:oxy_rec_sc"),
                       filter(minter_est2, term == "sub_div_f24:oxy_rec_sc")$estimate + estimate,
                       est2),

         est2 = ifelse(term %in% c("sub_div_f25:density_cod_sc",
                                   "sub_div_f26:density_cod_sc",
                                   "sub_div_f27:density_cod_sc",
                                   "sub_div_f28:density_cod_sc"),
                       filter(minter_est2, term == "sub_div_f24:density_cod_sc")$estimate + estimate,
                       est2),

         est2 = ifelse(term %in% c("sub_div_f25:density_saduria_rec_sc",
                                   "sub_div_f26:density_saduria_rec_sc",
                                   "sub_div_f27:density_saduria_rec_sc",
                                   "sub_div_f28:density_saduria_rec_sc"),
                       filter(minter_est2, term == "sub_div_f24:density_saduria_rec_sc")$estimate + estimate,
                       est2))

# upper
minter_est <- minter_est %>%
  mutate(upr2 = conf.high,
         upr2 = ifelse(term %in% c("sub_div_f25:biomass_spr_sc",
                                   "sub_div_f26:biomass_spr_sc",
                                   "sub_div_f27:biomass_spr_sc",
                                   "sub_div_f28:biomass_spr_sc"),
                       filter(minter_est2, term == "sub_div_f24:biomass_spr_sc")$conf.high + conf.high,
                       upr2),

         upr2 = ifelse(term %in% c("sub_div_f25:oxy_rec_sc",
                                   "sub_div_f26:oxy_rec_sc",
                                   "sub_div_f27:oxy_rec_sc",
                                   "sub_div_f28:oxy_rec_sc"),
                       filter(minter_est2, term == "sub_div_f24:oxy_rec_sc")$conf.high + conf.high,
                       upr2),

         upr2 = ifelse(term %in% c("sub_div_f25:density_cod_sc",
                                   "sub_div_f26:density_cod_sc",
                                   "sub_div_f27:density_cod_sc",
                                   "sub_div_f28:density_cod_sc"),
                       filter(minter_est2, term == "sub_div_f24:density_cod_sc")$conf.high + conf.high,
                       upr2),

         upr2 = ifelse(term %in% c("sub_div_f25:density_saduria_rec_sc",
                                   "sub_div_f26:density_saduria_rec_sc",
                                   "sub_div_f27:density_saduria_rec_sc",
                                   "sub_div_f28:density_saduria_rec_sc"),
                       filter(minter_est2, term == "sub_div_f24:density_saduria_rec_sc")$conf.high + conf.high,
                       upr2))


minter_est <- minter_est %>%
  mutate(lwr2 = conf.low,
         lwr2 = ifelse(term %in% c("sub_div_f25biomass_spr_sc",
                                   "sub_div_f26:biomass_spr_sc",
                                   "sub_div_f27:biomass_spr_sc",
                                   "sub_div_f28:biomass_spr_sc"),
                       filter(minter_est2, term == "sub_div_f24:biomass_spr_sc")$conf.low + conf.low,
                       lwr2),

         lwr2 = ifelse(term %in% c("sub_div_f25:oxy_rec_sc",
                                   "sub_div_f26:oxy_rec_sc",
                                   "sub_div_f27:oxy_rec_sc",
                                   "sub_div_f28:oxy_rec_sc"),
                       filter(minter_est2, term == "sub_div_f24:oxy_rec_sc")$conf.low + conf.low,
                       lwr2),

         lwr2 = ifelse(term %in% c("sub_div_f25:density_cod_sc",
                                   "sub_div_f26:density_cod_sc",
                                   "sub_div_f27:density_cod_sc",
                                   "sub_div_f28:density_cod_sc"),
                       filter(minter_est2, term == "sub_div_f24:density_cod_sc")$conf.low + conf.low,
                       lwr2),

         lwr2 = ifelse(term %in% c("sub_div_f25:density_saduria_rec_sc",
                                   "sub_div_f26:density_saduria_rec_sc",
                                   "sub_div_f27:density_saduria_rec_sc",
                                   "sub_div_f28:density_saduria_rec_sc"),
                       filter(minter_est2, term == "sub_div_f24:density_saduria_rec_sc")$conf.low + conf.low,
                       lwr2))

# minter_est <- minter_est %>% filter(!term %in% c("sub_div_f24:density_saduria_rec_sc",
#                                                  "sub_div_f24:oxy_rec_sc",
#                                                  "sub_div_f24:density_cod_sc",
#                                                  "biomass_spr_sc:sub_div_f24"))

p1 <- ggplot(minter_est %>% filter(group == "Saduria (rec)"), aes(term, est2)) +
  facet_wrap(~group) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray40", alpha = 0.5) +
  geom_errorbar(aes(ymin = lwr2, ymax = upr2), width = 0) +
  geom_point(size = 2.5) +
  labs(x = "", y = "Standardized coefficient") +
  theme_plot() +
  coord_flip() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(color = "none", fill = "none", shape = guide_legend(ncol = 4))

p2 <- ggplot(minter_est %>% filter(group == "Cod (rec)"), aes(term, est2)) +
  facet_wrap(~group) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray40", alpha = 0.5) +
  geom_errorbar(aes(ymin = lwr2, ymax = upr2), width = 0) +
  geom_point(size = 2.5) +
  labs(x = "", y = "Standardized coefficient") +
  theme_plot() +
  coord_flip() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(color = "none", fill = "none", shape = guide_legend(ncol = 4))

p3 <- ggplot(minter_est %>% filter(group == "Oxygen (rec)"), aes(term, est2)) +
  facet_wrap(~group) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray40", alpha = 0.5) +
  geom_errorbar(aes(ymin = lwr2, ymax = upr2), width = 0) +
  geom_point(size = 2.5) +
  labs(x = "", y = "Standardized coefficient") +
  theme_plot() +
  coord_flip() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(color = "none", fill = "none", shape = guide_legend(ncol = 4))

p4 <- ggplot(minter_est %>% filter(group == "Sprat (rec)"), aes(term, est2)) +
  facet_wrap(~group) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray40", alpha = 0.5) +
  geom_errorbar(aes(ymin = lwr2, ymax = upr2), width = 0) +
  geom_point(size = 2.5) +
  labs(x = "", y = "Standardized coefficient") +
  theme_plot() +
  coord_flip() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(color = "none", fill = "none", shape = guide_legend(ncol = 4))

p1 + p2 + p3 + p4
```
