---
title: "Condition model"
author: "Max Lindmark, Sean Andersson, Mayya Gogina and Michele Casini"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.asp = 0.618,
  fig.align ='center'
)
```

# Fit condition model with environmental and biological predictors
Fit main model (see exploratory scripts and model comparison), visualize results.

## Load packages

```{r packages, message=FALSE, warning=TRUE}
library(tidyverse); theme_set(theme_light(base_size = 12))
library(tidylog)
library(viridis)
library(marmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(RColorBrewer)
library(gganimate)
library(gifski)
library(latex2exp)
library(patchwork)
library(png)
library(RCurl)
library(wesanderson)
library(qwraps2) 
library(ggcorrplot)
library(ggridges)
#remotes::install_github("pbs-assess/sdmTMB", ref = "sim2")
library(sdmTMB)

# To load entire cache in interactive r session, do: 
# qwraps2::lazyload_cache_dir(path = "R/analysis/condition_model_cache/html")
```

## For maps

```{r read coastline data, message=FALSE, warning=FALSE}
# Specify map ranges
ymin = 52; ymax = 60; xmin = 10; xmax = 24

map_data <- rnaturalearth::ne_countries(
  scale = "medium",
  returnclass = "sf", continent = "europe")

# Crop the polygon for plotting and efficiency:
# st_bbox(map_data) # find the rough coordinates
swe_coast <- suppressWarnings(suppressMessages(
  st_crop(map_data,
          c(xmin = xmin, ymin = ymin, xmax = xmax, ymax = ymax))))

# Transform our map into UTM 33 coordinates, which is the equal-area projection we fit in:
utm_zone33 <- 32633
swe_coast_proj <- sf::st_transform(swe_coast, crs = utm_zone33)

# Define plotting theme for main plot
theme_plot <- function(base_size = 10, base_family = "") {
  theme_light(base_size = 10, base_family = "") +
    theme(
      axis.text.x = element_text(angle = 90),
      axis.text = element_text(size = 8),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 8),
      legend.position = "bottom",
      legend.key.height = unit(0.2, "cm"),
      legend.margin = margin(0, 0, 0, 0),
      legend.box.margin = margin(-5, -5, -5, -5),
      strip.text = element_text(size = 8, colour = 'gray10', margin = margin()),
      strip.background = element_rect(fill = "grey95")
      )
}

# Define plotting theme for facet_wrap map with years
theme_facet_map <- function(base_size = 10, base_family = "") {
  theme_light(base_size = 10, base_family = "") +
    theme(
        axis.text.x = element_text(angle = 90),
        axis.text = element_text(size = 6),
        strip.text = element_text(size = 8, colour = 'gray10', margin = margin()),
        strip.background = element_rect(fill = "gray95"),
        legend.position = c(0.7, 0.02),
        legend.direction = "horizontal"
      )
}

# Make default base map plot
xmin2 <- 303000; xmax2 <- 916000; xrange <- xmax2 - xmin2
ymin2 <- 5980000; ymax2 <- 6450000; yrange <- ymax2 - ymin2

plot_map_raster <- 
ggplot(swe_coast_proj) + 
  xlim(xmin2, xmax2) +
  ylim(ymin2, ymax2) +
  labs(x = "Longitude", y = "Latitude") +
  geom_sf(size = 0.3) + 
  theme_plot() +
  geom_sf() +
  NULL

plot_map_raster_labels <- 
  plot_map_raster + 
  annotate("text", label = "Sweden", x = xmin2 + 0.25*xrange, y = ymin2 + 0.8*yrange, color = "black", size = 1.9) +
  annotate("text", label = "Denmark", x = xmin2 + 0.029*xrange, y = ymin2 + 0.43*yrange, color = "black", size = 1.9, angle = 75) +
  annotate("text", label = "Germany", x = xmin2 + 0.07*xrange, y = ymin2 + 0.025*yrange, color = "black", size = 1.9) +
  annotate("text", label = "Poland", x = xmin2 + 0.55*xrange, y = ymin2 + 0.1*yrange, color = "black", size = 1.9) +
  annotate("text", label = "Russia", x = xmin2 + 0.95*xrange, y = ymin2 + 0.2*yrange, color = "black", size = 1.9) +
  annotate("text", label = "Lithuania", x = xmin2 + 1*xrange, y = ymin2 + 0.47*yrange, color = "black", size = 1.9, angle = 75) +
  annotate("text", label = "Latvia", x = xmin2 + 0.99*xrange, y = ymin2 + 0.68*yrange, color = "black", size = 1.9, angle = 75)
```

## Read data

```{r read and process data, message=FALSE, warning=FALSE}
# Read the split files and join them
d1 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/mdat_cond_(1_2).csv")
d2 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/mdat_cond_(2_2).csv")

d <- bind_rows(d1, d2)

# Calculate standardized variables
d <- d %>% 
  mutate(ln_length_cm = log(length_cm),
         ln_weight_g = log(weight_g),
         year_ct = year - mean(year),
         biomass_her_sc = biomass_her,
         biomass_her_sd_sc = biomass_her_sd,
         biomass_spr_sc = biomass_spr,
         biomass_spr_sd_sc = biomass_spr_sd,
         density_cod_sc = density_cod,
         density_cod_rec_sc = density_cod_rec,
         density_fle_sc = density_fle,
         density_fle_rec_sc = density_fle_rec,
         density_saduria_sc = density_saduria,
         density_saduria_rec_sc = density_saduria_rec,
         depth_sc = depth,
         depth_rec_sc = depth_rec,
         depth_sd_sc = depth_sd,
         oxy_sc = oxy,
         oxy_rec_sc = oxy_rec,
         oxy_sd_sc = oxy_sd,
         temp_sc = temp,
         temp_rec_sc = temp_rec,
         temp_sd_sc = temp_sd)  %>%
  mutate_at(c("biomass_her_sc", "biomass_her_sd_sc", "biomass_spr_sc", "biomass_spr_sd_sc",
              "density_cod_sc", "density_cod_rec_sc", 
              "density_fle_sc", "density_fle_rec_sc", 
              "density_saduria_sc", "density_saduria_rec_sc", 
              "depth_sc", "depth_rec_sc", "depth_sd_sc",
              "oxy_sc", "oxy_rec_sc", "oxy_sd_sc",
              "temp_sc", "temp_rec_sc", "temp_sd_sc"
              ),
            ~(scale(.) %>% as.vector)) %>% 
  mutate(year = as.integer(year))

# Plot correlation between variables
d_cor <- d %>% dplyr::select("biomass_her_sc", "biomass_her_sd_sc", "biomass_spr_sc", "biomass_spr_sd_sc",
                             "density_cod_sc", "density_cod_rec_sc", 
                             "density_fle_sc", "density_fle_rec_sc", 
                             "density_saduria_sc", "density_saduria_rec_sc", 
                             "depth_sc", "depth_rec_sc",
                             "oxy_sc", "oxy_rec_sc",
                             "temp_sc", "temp_rec_sc")

corr <- round(cor(d_cor), 1)

ggcorrplot(corr, hc.order = TRUE, type = "lower", lab = TRUE, lab_size = 2.5) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3))

ggsave("figures/supp/condition/correlation_variables.png", width = 6.5, height = 6.5, dpi = 600)
```

## Read the prediction grids

```{r read and process prediction grid, message=FALSE, warning=FALSE}
pred_grid1 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/pred_grid_(1_2).csv")
pred_grid2 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/pred_grid_(2_2).csv")

pred_grid <- bind_rows(pred_grid1, pred_grid2)

pred_grid <- pred_grid %>% mutate(year = as.integer(year))

# Scale the variables with respect to data! First calculate means in data
data_means <- d %>%
  dplyr::select(biomass_her, biomass_her_sd, biomass_spr, biomass_spr_sd,
                density_cod, density_cod_rec, 
                density_fle, density_fle_rec, 
                density_saduria, density_saduria_rec, 
                depth, depth_rec, depth_sd,
                oxy, oxy_rec, oxy_sd,
                temp, temp_rec, temp_sd) %>%
  mutate_all(~mean(.)) %>%
  distinct(.keep_all = TRUE)

data_stdev <- d %>%
  dplyr::select(biomass_her, biomass_her_sd, biomass_spr, biomass_spr_sd,
                density_cod, density_cod_rec, 
                density_fle, density_fle_rec, 
                density_saduria, density_saduria_rec, 
                depth, depth_rec, 
                oxy, oxy_rec, 
                temp, temp_rec
                ) %>%
  mutate_all(~sd(.)) %>%
  distinct(.keep_all = TRUE)

# Before actually scaling, replace the ices-rectangle pelagic values that are NA with the mean across rectangles 
# in the sub division. Replace the sub-division pelagic values that are NA with the mean in the year.
# Note that the sub-division values are not the sum of the rectangles (due to missing rectangles), so 
# I need to calculate a sub-division mean across rectangles within each sub-division
pred_grid <- pred_grid %>% 
  group_by(year) %>% 
  mutate(median_biomass_sprat_across_sub_div = median(biomass_spr_sd, na.rm = TRUE),
         median_biomass_herring_across_sub_div = median(biomass_her_sd, na.rm = TRUE)) %>% 
  ungroup() %>% # Replace sub_divsion NA's with the median across sub_divisions in that year
  mutate(biomass_spr_sd = ifelse(is.na(biomass_spr_sd) == TRUE, median_biomass_sprat_across_sub_div, biomass_spr_sd),
         biomass_her_sd = ifelse(is.na(biomass_her_sd) == TRUE, median_biomass_herring_across_sub_div, biomass_her_sd)) %>% 
  group_by(year, sub_div) %>% 
  mutate(median_biomass_sprat_across_rect_in_sub_div = median(biomass_spr, na.rm = TRUE),
         median_biomass_herring_across_rect_in_sub_div = median(biomass_her, na.rm = TRUE)) %>% 
  ungroup() %>% # Replace rectangle NA's with the median across rectangles in that year and sub-division
  mutate(biomass_spr = ifelse(is.na(biomass_spr) == TRUE, median_biomass_sprat_across_rect_in_sub_div, biomass_spr),
         biomass_her = ifelse(is.na(biomass_her) == TRUE, median_biomass_herring_across_rect_in_sub_div, biomass_her)) %>% 
  # Since I still have some NAs (some sub-divisions do not have a single rectangle in some years), I will will those rectangles 
  # with the sub-division value divided by the number of rectangles in that sub division
  group_by(year, sub_div) %>% 
  mutate(n_rect = length(unique(ices_rect))) %>% 
  ungroup() %>% 
  mutate(biomass_spr = ifelse(is.na(biomass_spr) == TRUE, biomass_spr_sd/n_rect, biomass_spr),
         biomass_her = ifelse(is.na(biomass_her) == TRUE, biomass_her_sd/n_rect, biomass_her))

pred_grid <- pred_grid %>%
  mutate(ln_length_cm = log(1),
         year_ct = year - mean(year),
         biomass_her_sc = (biomass_her - data_means$biomass_her) / data_stdev$biomass_her,
         biomass_her_sd_sc = (biomass_her_sd - data_means$biomass_her_sd) / data_stdev$biomass_her_sd,
         biomass_spr_sc = (biomass_spr - data_means$biomass_spr) / data_stdev$biomass_spr,
         biomass_spr_sd_sc = (biomass_spr_sd - data_means$biomass_spr_sd) / data_stdev$biomass_spr_sd,
         density_cod_sc = (density_cod - data_means$density_cod) / data_stdev$density_cod,
         density_cod_rec_sc = (density_cod_rec - data_means$density_cod_rec) / data_stdev$density_cod_rec,
         density_fle_sc = (density_fle - data_means$density_fle) / data_stdev$density_fle,
         density_fle_rec_sc = (density_fle_rec - data_means$density_fle_rec) / data_stdev$density_fle_rec,
         density_saduria_sc = (density_saduria - data_means$density_saduria) / data_stdev$density_saduria,
         density_saduria_rec_sc = (density_saduria_rec - data_means$density_saduria_rec) / data_stdev$density_saduria_rec,
         depth_sc = (depth - data_means$depth) / data_stdev$depth,
         depth_rec_sc = (depth_rec - data_means$depth_rec) / data_stdev$depth_rec,
         oxy_sc = (oxy - data_means$oxy) / data_stdev$oxy,
         oxy_rec_sc = (oxy_rec - data_means$oxy_rec) / data_stdev$oxy_rec,
         temp_sc = (temp - data_means$temp) / data_stdev$temp,
         temp_rec_sc = (temp_rec - data_means$temp_rec) / data_stdev$temp_rec,
         )

# Plot on map:
ggplot(pred_grid2, aes(X, Y)) + geom_point(size = 0.1) + facet_wrap(~year)

pred_grid %>% 
  drop_na() %>% 
  ggplot(., aes(X, Y)) + geom_point(size = 0.1) + facet_wrap(~year)

pred_grid %>% 
  drop_na(biomass_spr_sd) %>% 
  ggplot(., aes(X, Y)) + geom_point(size = 0.1) + facet_wrap(~year)

unique(is.na(pred_grid))

pred_grid %>% 
  drop_na(biomass_spr_sc, biomass_spr_sd_sc, biomass_her_sc, biomass_her_sd_sc) %>% 
  ggplot(., aes(X, Y)) + geom_point(size = 0.1) + facet_wrap(~year)
```

```{r, include=FALSE}
# Include some tests for the manuscript
# pred_grid_test <- pred_grid %>%
#   mutate(oxy_cat = ifelse(oxy > 4.3, "normal", "low"),
#          oxy_cat = ifelse(oxy < 1, "dead", oxy_cat)) %>%
#   filter(year %in% c(1993, 2018))
#  
# plot_map_raster +
#   geom_raster(data = pred_grid_test,
#               aes(x = X * 1000, y = Y * 1000, fill = oxy_cat)) +
#   facet_wrap(~ year) +
#   scale_fill_manual(values = c("black", "gray70", "skyblue")) +
#   theme_light() +
#   geom_sf() + 
#   NULL

# ggplot(pred_grid, aes(depth, oxy)) + geom_point() +
#   geom_hline(yintercept = 1, col = "red") +
#   geom_hline(yintercept = 4.3, col = "red")
# 
# m1 <- lm(oxy ~ depth, data = pred_grid)
# nd <- data.frame(depth = c(40, 45, 50, 60, 70))
# nd$pred <- predict(m1, newdata = nd)
# nd
# nd$pred[2] - nd$pred[1]
# 
# plot(temp ~ depth, data = pred_grid)
# m1 <- lm(temp ~ depth, data = filter(pred_grid, depth < 100))
# nd <- data.frame(depth = c(40, 45))
# nd$pred <- predict(m1, newdata = nd)
# nd
# nd$pred[2] - nd$pred[1]
# 
# pred_grid %>% 
#   group_by(sub_div, year) %>% 
#   summarise(mean_oxy = mean(oxy)) %>% 
#   ungroup() %>% 
#   ggplot(., aes(year, mean_oxy, color = sub_div)) +
#   geom_line()
```

## Make spde mesh

```{r make mesh, results='hide', message=FALSE}
spde <- make_mesh(d, xy_cols = c("X", "Y"),
                  n_knots = 200,  
                  type = "kmeans", seed = 42)

# Plot and save spde
png(file = "figures/supp/condition/spde.png", units = "in", width = 6.5, height = 6.5, res = 300)
plot(spde)
dev.off()
```

## Fit model

```{r fit full model, cache=TRUE}
ptm <- proc.time()
mfull <- sdmTMB(
  formula = ln_weight_g ~ -1 + ln_length_cm + biomass_her_sc + biomass_her_sd_sc +
    biomass_spr_sc + biomass_spr_sd_sc +
    density_cod_sc + density_cod_rec_sc + #density_cod_sd_sc +
    density_fle_sc + density_fle_rec_sc + #density_fle_sd_sc +
    density_saduria_sc + density_saduria_rec_sc + #density_saduria_sd_sc + 
    depth_sc + depth_rec_sc + #depth_sd_sc +
    oxy_sc + oxy_rec_sc + #oxy_sd_sc +
    temp_sc + temp_rec_sc, #+ #temp_sd_sc,
  time_varying = ~ 1, data = d, time = "year", mesh = spde,
  family = student(link = "identity", df = 5),
  spatiotemporal = "AR1", spatial = "on", silent = TRUE, reml = FALSE,
  control = sdmTMBcontrol(newton_loops = 1, eval.max = 10000, iter.max = 10000, nlminb_loops = 2))
proc.time() - ptm

tidy(mfull, conf.int = TRUE) %>% arrange(desc(abs(estimate)))
tidy(mfull, effects = "ran_par", conf.int = TRUE) 
```

## Plot residuals

```{r extract residuals}
d$residuals_mfull <- residuals(mfull)

png(file = "figures/supp/condition/qq.png", units = "in", width = 6.5, height = 6.5, res = 300)
qqnorm(d$residuals_mfull); abline(a = 0, b = 1)
dev.off()

# Residuals on map
plot_map_raster +
  geom_point(data = d, aes(x = X * 1000, y = Y * 1000, color = residuals_mfull)) +
  scale_colour_gradient2() +
  facet_wrap(~ year, ncol = 5) +
  labs(color = "residuals") +
  geom_sf() +
  NULL

ggsave("figures/supp/condition/spatial_resid.png", width = 6.5, height = 6.5, dpi = 600)

# Residuals vs length
# ggplot(d, aes(ln_length_cm, residuals_mfull)) +
#   geom_point(alpha = 0.1, color = "grey20") + 
#   geom_abline(color = "red", slope = 0, linetype = 2) +
#   geom_smooth()
# 
# ggsave("figures/supp/condition/residuals_vs_length.png", width = 6.5, height = 6.5, dpi = 600)

# Residuals vs length and year
ggplot(d, aes(x = ln_length_cm, y = residuals_mfull, color = residuals_mfull)) +
  geom_point(size = 0.1, alpha = 1) + 
  geom_hline(yintercept = 0, size = 0.5, color = "black", linetype = 2, alpha = 0.5) +
  labs(x = "log(length)", y = "Residuals", color = "") +
  stat_smooth(color = "black", size = 0.5) +
  facet_wrap(~year, ncol = 5) +
  scale_color_gradient2() +
  theme_facet_map()

ggsave("figures/supp/condition/residuals_vs_length_and_year.png", width = 6.5, height = 6.5, dpi = 600)

# ggplot(d, aes(year, ln_length_cm, z = residuals_mfull)) +
#   stat_summary_2d(bins = 40) +
#   scale_fill_gradient2()
# 
# ggsave("figures/supp/condition/residuals_vs_length_and_year2.png", width = 6.5, height = 6.5, dpi = 600)
```

## Annual condition factor

```{r annual condition factor, message=FALSE,cache=TRUE}
# From these models, predict annual condition factor
# I will use the prediction grid with ALL covariates set to 0, incl. depth, temp and oxygen
# Grabbing the number of cells to help with calculating the average
# ncells <- filter(pred_grid, year == max(pred_grid$year)) %>% nrow()
# 
# # Use the `area` argument here to turn the total into an average by giving it one over the number of cells
# pred_avg <- predict(mfull, newdata = pred_grid, return_tmb_object = TRUE, area = 1/ncells)
# 
# # Make a little helper function... bias correction shouldn't do anything here because of the identity link
# get_average_condition <- function(obj, level = 0.95, ...)  {
#   sdmTMB:::get_generic(obj, value_name = "link_total",
#                        bias_correct = FALSE, level = level, trans = I, ...)
# }
# 
# avg_pred <- get_average_condition(pred_avg)
```

## Annual condition factor by sub-division

```{r annual condition factor by sub-division, message=FALSE,cache=TRUE}
# # From these models, predict annual condition factor
# # I will use the prediction grid with ALL covariates set to 0, incl. depth, temp and oxygen
# # Grabbing the number of cells to help with calculating the average
# ncells24 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 24) %>% nrow()
# ncells25 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 25) %>% nrow()
# ncells26 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 26) %>% nrow()
# ncells27 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 27) %>% nrow()
# ncells28 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 28) %>% nrow()
# 
# # Use the `area` argument here to turn the total into an average by giving it one over the number of cells
# pred_avg_24 <- predict(mfull, newdata = filter(pred_grid, sub_div == 24), return_tmb_object = TRUE, area = 1/ncells24)
# pred_avg_25 <- predict(mfull, newdata = filter(pred_grid, sub_div == 25), return_tmb_object = TRUE, area = 1/ncells25)
# pred_avg_26 <- predict(mfull, newdata = filter(pred_grid, sub_div == 26), return_tmb_object = TRUE, area = 1/ncells26)
# pred_avg_27 <- predict(mfull, newdata = filter(pred_grid, sub_div == 27), return_tmb_object = TRUE, area = 1/ncells27)
# pred_avg_28 <- predict(mfull, newdata = filter(pred_grid, sub_div == 28), return_tmb_object = TRUE, area = 1/ncells28)
# 
# avg_pred_24 <- get_average_condition(pred_avg_24)
# avg_pred_25 <- get_average_condition(pred_avg_25)
# avg_pred_26 <- get_average_condition(pred_avg_26)
# avg_pred_27 <- get_average_condition(pred_avg_27)
# avg_pred_28 <- get_average_condition(pred_avg_28)
# 
# # Combine and plot
# avg_pred_24 <- avg_pred_24 %>% mutate(sub_div = 24)
# avg_pred_25 <- avg_pred_25 %>% mutate(sub_div = 25)
# avg_pred_26 <- avg_pred_26 %>% mutate(sub_div = 26)
# avg_pred_27 <- avg_pred_27 %>% mutate(sub_div = 27)
# avg_pred_28 <- avg_pred_28 %>% mutate(sub_div = 28)
# 
# div_index <- bind_rows(avg_pred_24, avg_pred_25, avg_pred_26, avg_pred_27, avg_pred_28)
```

## Annual indices using `get_index_sims` method to avoid slow bias correction (not done above!)

```{r sim version condition index, cache=TRUE}
ncells <- filter(pred_grid, year == max(pred_grid$year)) %>% nrow()

pred_avg_sim <- predict(mfull, newdata = pred_grid, sims = 100)

index_avg_sim <- get_index_sims(pred_avg_sim, area = rep(1/ncells, nrow(pred_avg_sim)))
index_avg_sims <- get_index_sims(pred_avg_sim, area = rep(1/ncells, nrow(pred_avg_sim)),
                                 return_sims = TRUE) # This is for calculation on samples!

ggplot(index_avg_sim, aes(year, log(est), ymin = log(lwr), ymax = log(upr))) +
  geom_line() +
  geom_ribbon(alpha = 0.4)
```

## Difference in log(condition factor) between the end and start of time period using sims

```{r simulate log condition factor,cache=TRUE}
colnames(pred_grid)

# Plot the sims
ggplot(index_avg_sims, aes(y = as.factor(year), x = log(.value))) + 
  geom_density_ridges_gradient(rel_min_height = 0.01, alpha = 0.8, color = "grey10", size = 0.2,
                               quantile_lines = TRUE, quantiles = 0.5) + 
  coord_flip()

index_avg_sims %>% 
  filter(year %in% c(1993, 2008)) %>% 
  ggplot(., aes(as.factor(year), .value)) +
  geom_violin(width = 1.3)

# Now reshape data again to calculate row-wise differences
year_diff <- index_avg_sims %>% 
  filter(year %in% c(1993, 2008)) %>% 
  mutate(.value = log(.value)) %>% 
  pivot_wider(names_from = year, values_from = .value) %>% 
  mutate("2008-1993" = `2008` - `1993`) %>% 
  pivot_longer(2:4) %>% 
  ungroup()

# Plot these
p1 <- ggplot(year_diff) +
  geom_density(data = filter(year_diff, !name == "2008-1993"),
               aes(value, fill = name), alpha = 0.6) +
  scale_fill_brewer(palette = "Dark2") +
  coord_cartesian(expand = 0)

p2 <- ggplot(year_diff) +
  geom_density(data = filter(year_diff, name == "2008-1993"),
               aes(value), alpha = 0.6, fill = "grey") +
  geom_vline(xintercept = 0, linetype = 2) 

p1/p2

# Calculate quantiles for each level (to go in the results section)
year_diff %>% 
  group_by(name) %>% 
  summarise(median = quantile(value, probs = 0.5),
            upr97.5 = quantile(value, probs = 0.975),
            lwr2.5 = quantile(value, probs = 0.025))

# Percent change in condition factor
year_diff %>% 
  group_by(name) %>% 
  pivot_wider(names_from = name, values_from = value) %>%
  mutate(perc_change = (`2008`-`1993`)/`1993`*100) %>% 
  ungroup() %>% 
  summarise(lwr2.5 = quantile(perc_change, probs = 0.025),
            median = quantile(perc_change, probs = 0.5),
            upr97.5 = quantile(perc_change, probs = 0.975))

# Percent change in weight
b <- mfull$sd_report$par.fixed[1]
  
year_diff %>% 
  group_by(name) %>% 
  pivot_wider(names_from = name, values_from = value) %>%
  mutate(w1 = exp(`1993`)*30^b, # using length = 30 here
         w2 = exp(`2008`)*30^b, 
         perc_change = (`w2`-`w1`)/`w1`*100) %>% 
  ungroup() %>% 
  summarise(lwr2.5 = quantile(perc_change, probs = 0.025),
            median = quantile(perc_change, probs = 0.5),
            upr97.5 = quantile(perc_change, probs = 0.975))
```

## Simulated annual condition factor by sub-division 

```{r sim annual condition factor by sub-division, message=FALSE,cache=TRUE}
# From these models, predict annual condition factor
# I will use the prediction grid with ALL covariates set to 0, incl. depth, temp and oxygen
# Grabbing the number of cells to help with calculating the average
ncells24 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 24) %>% nrow()
ncells25 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 25) %>% nrow()
ncells26 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 26) %>% nrow()
ncells27 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 27) %>% nrow()
ncells28 <- filter(pred_grid, year == max(pred_grid$year) & sub_div == 28) %>% nrow()

# Use the `area` argument here to turn the total into an average by giving it one over the number of cells
pred_avg_24_sim <- predict(mfull, newdata = filter(pred_grid, sub_div == 24), sims = 100)
pred_avg_25_sim <- predict(mfull, newdata = filter(pred_grid, sub_div == 25), sims = 100)
pred_avg_26_sim <- predict(mfull, newdata = filter(pred_grid, sub_div == 26), sims = 100)
pred_avg_27_sim <- predict(mfull, newdata = filter(pred_grid, sub_div == 27), sims = 100)
pred_avg_28_sim <- predict(mfull, newdata = filter(pred_grid, sub_div == 28), sims = 100)

index_avg_24_sim <- get_index_sims(pred_avg_24_sim, area = rep(1/ncells24, nrow(pred_avg_24_sim)))
index_avg_25_sim <- get_index_sims(pred_avg_25_sim, area = rep(1/ncells25, nrow(pred_avg_25_sim)))
index_avg_26_sim <- get_index_sims(pred_avg_26_sim, area = rep(1/ncells26, nrow(pred_avg_26_sim)))
index_avg_27_sim <- get_index_sims(pred_avg_27_sim, area = rep(1/ncells27, nrow(pred_avg_27_sim)))
index_avg_28_sim <- get_index_sims(pred_avg_28_sim, area = rep(1/ncells28, nrow(pred_avg_28_sim)))

# Combine and plot
index_avg_24_sim <- index_avg_24_sim %>% mutate(sub_div = 24)
index_avg_25_sim <- index_avg_25_sim %>% mutate(sub_div = 25)
index_avg_26_sim <- index_avg_26_sim %>% mutate(sub_div = 26)
index_avg_27_sim <- index_avg_27_sim %>% mutate(sub_div = 27)
index_avg_28_sim <- index_avg_28_sim %>% mutate(sub_div = 28)

div_index <- bind_rows(index_avg_24_sim, index_avg_25_sim, index_avg_26_sim, index_avg_27_sim, index_avg_28_sim)
```

## Condition factor, condition vs length, quantiles and size-distribution and cv

```{r plot condition index and condition factor, message=FALSE}
# Add in the Le Cren condition index for d
mean_tw_intercept <- mean(mfull$sd_report$par.random[1:27])
d$le_cren <- d$weight_g / (exp(mean_tw_intercept)*d$length_cm^2.98)

div_index2 <- bind_rows(mutate(div_index, sub_div = as.factor(sub_div)),
                        mutate(index_avg_sim, sub_div = as.factor("Total")))

pal <- c(brewer.pal(n = 5, name = "Dark2"), "grey10")

# Remember to log now that we use sims
cond_index <- ggplot(div_index2, aes(year, log(est), color = sub_div, linetype = sub_div)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(y = "log(condition factor)", x = "Year") +
  geom_line(alpha = 0.8) +
  geom_point(alpha = 0) + 
  scale_color_manual(values = pal, name = "ICES\nSub division") +
  scale_linetype_manual(values = c(1,1,1,1,1,0)) +
  geom_point(data = index_avg_sim, aes(year, log(est)), inherit.aes = FALSE, 
             size = 2.3, shape = 21, fill = "gray1", color = "white") +
  geom_errorbar(data = index_avg_sim, inherit.aes = FALSE, 
                aes(x = year, ymax = log(upr), ymin = log(lwr)),
                width = 0.6, alpha = 0.8, color = "gray1", size = 0.5) +
  theme(axis.title = element_text(size = 9),
        legend.position = c(0.85, 0.75),
        legend.background = element_blank(),
        legend.title = element_text(size = 9),
        legend.spacing.x = unit(0, 'cm'),
        plot.margin = unit(c(0.4, 0.4, 0.4, 0), "cm")) +
  guides(linetype = "none",
         color = guide_legend(keywidth = 0.1, keyheight = 0.1, default.unit = "inch",
                              override.aes = list(linetype = c(1,1,1,1,1,0),
                                                  shape = c(NA,NA,NA,NA,NA,16),
                                                  alpha = rep(0.8, 6),
                                                  color = pal))) + 
  NULL

cond_index

# Plot change in condition over size
# Plot the slopes
cond_year_slopes <- d %>%
  mutate(lenCls = cut(length_cm, breaks = seq(0, 150, 10)),
         lenCls_fct = factor(lenCls)) %>% 
  mutate(year_ct = year - 1993) %>% 
  split(.$lenCls_fct) %>%
  purrr::map(~lm(le_cren ~ year_ct, data = .x)) %>%
  purrr::map_df(broom::tidy, .id = 'lenCls_fct') %>%
  filter(term %in% c('year_ct', "(Intercept)")) %>% 
  dplyr::select(-c(statistic, p.value)) %>% 
  filter(term == "year_ct")

cond_slope_fig <- cond_year_slopes %>% 
  separate(lenCls_fct, c("lwr", "upr"), sep = ",") %>%
  separate(upr, c("upr", "scrap"), sep = "]") %>%
  dplyr::select(-scrap) %>% 
  mutate(upr_num = as.numeric(upr)) %>% 
  filter(upr_num > 10 & upr_num < 80) %>% 
  ggplot(., aes(as.factor(upr_num), estimate)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = 2, color = "red") +
  geom_point(size = 1.2, position = position_dodge(width = 0.5), color = "gray40") + 
  geom_errorbar(aes(x = as.factor(upr_num),
                    ymax = estimate + 1.96*std.error,
                    ymin = estimate - 1.96*std.error),
                width = 0.3, position = position_dodge(width = 0.5), color = "gray40") +
  labs(y = "Slope (condition index ~ year)", x = "Length-class (upper value) [cm]") +
  theme(axis.title = element_text(size = 9)) +
  NULL

size_dist <- ggplot(d, aes(length_cm)) +
  geom_density(alpha = 0.8, fill = "gray70", color = "gray50") +
  coord_cartesian(expand = 0, xlim = c(14, 76), ylim = c(0, 0.031)) + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 7)) +
  labs(y = "Density", x = "Length-class (upper value) [cm]") +
  theme(axis.title = element_text(size = 9)) + 
  NULL

cond_slope_fig / size_dist + plot_layout(heights = c(3, 1))

# Plot deciles
pal <- brewer.pal(n = 8, name = "Dark2")

deciles_sum <- d %>% 
  group_by(year) %>% 
  summarise("1st" = hutils::weighted_quantile(v = le_cren, p = c(0.1)),
            "5th (median)" = hutils::weighted_quantile(v = le_cren, p = c(0.5)),
            "9th" = hutils::weighted_quantile(v = le_cren, p = c(0.9))) %>% 
  ungroup() %>% 
  pivot_longer(2:4, names_to = "Deciles", values_to = "le_cren") %>% 
  group_by(Deciles)

deciles <- ggplot(deciles_sum, aes(year, le_cren, color = factor(Deciles), fill = factor(Deciles))) +
  stat_smooth(method = "gam", formula = y~s(x, k = 3),
              se = TRUE, show.legend = FALSE) + 
  stat_smooth(method = "gam", formula = y~s(x, k = 3), 
              se = FALSE) + 
  geom_point() +
  labs(x = "Year", y = "Le Cren's condition index", color = "Deciles") +
  scale_color_manual(values = rev(pal)) +
  scale_fill_manual(values = rev(pal)) +
  guides(color = guide_legend(override.aes = list(fill = NA))) + 
  theme() +
  theme(axis.title = element_text(size = 9),
        legend.position = c(0.85, 0.85),
        legend.background = element_blank(),
        legend.title = element_text(size = 9),
        legend.spacing.x = unit(0, 'cm'),
        plot.margin = unit(c(0.4, 0.4, 0.4, 0), "cm")) +
  guides(fill = "none",
         color = guide_legend(keywidth = 0.1,
                              keyheight = 0.1, 
                              default.unit = "inch")) + 
  NULL
  
cond_index + cond_slope_fig + deciles + size_dist +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2, widths = c(2, 1))

ggsave("figures/cond_index_sizes_change_deciles.png", width = 6.5, height = 6.5, dpi = 600)
```

## Effect sizes

```{r extract coefficients, message=FALSE}
# Extract random and fixed coefficients from the full model
mfull_est <- bind_rows(tidy(mfull, effects = "ran_par", conf.int = TRUE) %>%
                         filter(term %in% c("sigma_O", "sigma_E")),
                       tidy(mfull, effects = "fixed", conf.int = TRUE) %>%
                         filter(!term %in% c("ln_length_cm"))) %>%
  mutate(term = factor(term))

mfull_est <- mfull_est %>% 
  mutate(group = "Herring",
         group = ifelse(term %in% c("biomass_spr_sc", "biomass_spr_sd_sc"), "Sprat", group),
         group = ifelse(term %in% c("density_cod_rec_sc", "density_cod_sc"), "Cod", group),
         group = ifelse(term %in% c("density_fle_rec_sc", "density_fle_sc"), "Flounder", group),
         group = ifelse(term %in% c("density_saduria_rec_sc", "density_saduria_sc"), "Saduria", group),
         group = ifelse(term %in% c("depth_rec_sc", "depth_sc"), "Depth", group),
         group = ifelse(term %in% c("oxy_rec_sc", "oxy_sc"), "Oxygen", group),
         group = ifelse(term %in% c("temp_rec_sc", "temp_sc"), "Temp", group),
         group = ifelse(term == "sigma_E", "Random", group),
         group = ifelse(term == "sigma_O", "Random", group)
         )

mfull_est <- mfull_est %>% 
  mutate(scale = "Sub-division",
         scale = ifelse(term %in% c("biomass_her_sc", "biomass_spr_sc", "density_cod_rec_sc", "density_fle_rec_sc",
                                    "density_saduria_rec_sc", "depth_rec_sc", "oxy_rec_sc", "temp_rec_sc"), "Rectangle", scale),
         scale = ifelse(term %in% c("density_cod_sc", "density_fle_sc", "density_saduria_sc",
                                    "depth_sc", "oxy_sc", "temp_sc"), "Haul", scale),
         scale = ifelse(term == "sigma_E", "Spatial/spatiotemporal s.d.", scale),
         scale = ifelse(term == "sigma_O", "Spatial/spatiotemporal s.d.", scale)
         )

# Plot effects
# This is the order changed labels should go in!
levels(mfull_est$term)

# I want the random effects to be dark gray...
#pal2 <- c(pal[1:3], "black", pal[4:8])
pal <- brewer.pal(n = 8, name = "Dark2")
pal2 <- c(pal[1:5], "black", pal[6:8])

# Sort the terms so that random effects are at the top...
mfull_est <- mfull_est %>% 
  mutate(term2 = ifelse(term %in% c("sigma_E", "sigma_O"), 2, 1))

#ggplot(mfull_est, aes(reorder(term, estimate), estimate)) +
ggplot(mfull_est, aes(reorder(term, term2), estimate, color = group, fill = group, shape = scale)) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray40", alpha = 0.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_point(size = 2.5) +
  scale_color_manual(values = pal2, name = "") +
  scale_shape_manual(values = c(15, 19, 21, 17), name = "") +
  scale_fill_manual(values = rep("white", 9), name = "") +
  labs(x = "", y = "Standardized coefficient") +
  scale_x_discrete(breaks = levels(mfull_est$term),
                   labels = c(expression(Herring[rec]),
                              expression(Herring[sub]),
                              expression(Sprat[rec]),
                              expression(Sprat[sub]),
                              expression(Cod[rec]),
                              expression(Cod[haul]),
                              expression(Flounder[rec]),
                              expression(Flounder[haul]),
                              expression(Saduria~entomon[rec]),
                              expression(Saduria~entomon[haul]),
                              expression(Depth[rec]),
                              expression(Depth[haul]),
                              expression(Oxygen[rec]),
                              expression(Oxygen[haul]),
                              expression(sigma[E]),
                              expression(sigma[O]),
                              expression(Temp[rec]),
                              expression(Temp[haul])
                              )) +
  coord_flip() +
  theme(plot.margin = unit(c(0.4, 0.4, 0.4, 0), "cm"),
        legend.position = "bottom") +
  guides(color = "none", fill = "none", shape = guide_legend(ncol = 2)) + 
  NULL
  
ggsave("figures/effect_size.png", width = 4.5, height = 6.5, dpi = 600)

# Just a test to see the labels were alright
#ggplot(mfull_est, aes(reorder(term, estimate), estimate)) +
ggplot(mfull_est, aes(term, estimate)) +
  geom_hline(yintercept = 0, linetype = 2, color = "red", alpha = 0.5) +
  geom_point(size = 2) +
  labs(x = "", y = "Standardized coefficient") +
  coord_flip()

# Now save this for comparison in sensitivity analysis script
write.csv(mfull_est, "output/mfull_est.csv")
```

## Map-prediction
  
```{r predict on grid with covariates, message=FALSE, fig.width=12}
pred_map <- predict(mfull, newdata = pred_grid)

pal <- wes_palette("Zissou1", 21, type = "continuous")

summary(pred_map$est)
hist(pred_map$est)

# For plotting, remove estimates larger than the 99.99th percentile and smaller than the 0.01th percentile
pred_map2 <- pred_map %>% 
  mutate(lwr = hutils::weighted_quantile(v = est, p = c(0.0001)),
         upr = hutils::weighted_quantile(v = est, p = c(0.9999))) %>% 
  filter(est > lwr & est < upr)

plot_map_raster +
  geom_raster(data = pred_map2, aes(x = X*1000, y = Y*1000, fill = est)) +
  scale_fill_gradientn(colours = rev(pal), name = "log(condition factor)") +
  facet_wrap(~ year, ncol = 5) +
  theme_facet_map() +
  geom_sf() +
  NULL

ggsave("figures/supp/condition/all_years_condition_map_covars.png", width = 6.5, height = 6.5, dpi = 600)

# And a smaller map for selected years
plot_map_raster_labels +
  geom_raster(data = filter(pred_map2, year %in% c(1994, 2001, 2008, 2018)),
              aes(x = X*1000, y = Y*1000, fill = est)) +
  scale_fill_gradientn(colours = rev(pal), name = "log(condition factor)") + 
  facet_wrap(~ year, ncol = 2) +
  theme_plot() +
  geom_sf() +
  NULL

ggsave("figures/condition_map_covars.png", width = 6.5, height = 6.5, dpi = 600)
```

```{r gganimate condition,cache=TRUE,message=FALSE}
# p <- plot_map_raster +
#   geom_raster(data = pred_map2, aes(x = X*1000, y = Y*1000, fill = est)) +
#   scale_fill_gradientn(colours = rev(pal), name = "log(condition factor)") +
#   theme_plot() +
#   geom_sf() +
#   NULL
# 
# # Here comes the gganimate specific bits
# anim <- p +
#   labs(title = 'Year: {frame_time}') +
#   transition_time(as.integer(year)) +
#   ease_aes('linear')
# 
# gganimate::animate(anim, height = 600, width = 600)
# 
# anim_save(filename = "/Users/maxlindmark/Desktop/R_STUDIO_PROJECTS/cod_condition/output/gif/cond_oxy.gif")
```

## Plot spatial and spatiotemporal random effects:
  
```{r spatial omega map, message=FALSE, fig.width=12}
# Plot spatiotemporal random effect
plot_map_raster +
  geom_raster(data = pred_map, aes(x = X * 1000, y = Y * 1000, fill = epsilon_st)) +
  scale_fill_gradient2() +
  facet_wrap(~ year, ncol = 5) +
  ggtitle("Spatiotemporal random effects") +
  geom_sf() +
  NULL

ggsave("figures/supp/condition/epsilon_st_map.png", width = 6.5, height = 6.5, dpi = 600)

# Plot spatial random effect
plot_map_raster +
  geom_raster(data = filter(pred_map, year == 1999), aes(x = X * 1000, y = Y * 1000, fill = omega_s)) +
  scale_fill_gradient2() +
  facet_wrap(~ year, ncol = 5) +
  theme_plot() + 
  ggtitle("Spatial random effects") +
  geom_sf() +
  NULL

ggsave("figures/supp/condition/omega_s_map.png", width = 6.5, height = 6.5, dpi = 600)
```

## Calculate the "spatial trend" from the estimates:
  
```{r calculate "spatial trend", message=FALSE}
# Fit a linear model to each prediction grid of the estimate over time
# https://community.rstudio.com/t/extract-slopes-by-group-broom-dplyr/2751/7
time_slopes_by_year <- pred_map %>%
  mutate(id = paste(X, Y, sep = "_")) %>%
  split(.$id) %>%
  purrr::map(~lm(est ~ year, data = .x)) %>%
  purrr::map_df(broom::tidy, .id = 'id') %>%
  filter(term == 'year')

# Plot the slopes
time_slopes_by_year2 <- time_slopes_by_year %>%
  separate(id, c("X", "Y"), sep = "_") %>%
  mutate(X = as.numeric(X),
         Y = as.numeric(Y))

plot_map_raster +
  geom_raster(data = time_slopes_by_year2, aes(x = X * 1000, y = Y * 1000, fill = estimate)) +
  scale_fill_gradient2(midpoint = 0) +
  facet_wrap(~ term, ncol = 5) +
  theme_plot() +
  ggtitle("Time slopes by each pred grid") +
  geom_sf() +
  NULL

ggsave("figures/supp/condition/spatial_trend_condition.png", width = 6.5, height = 6.5, dpi = 600)

# Plot the upper CI
# ggplot(swe_coast_proj) +
#   geom_raster(data = time_slopes_by_year2, aes(x = X * 1000, y = Y * 1000, fill = estimate + std.error*1.96)) +
#   geom_sf(size = 0.3) +
#   scale_fill_gradient2(midpoint = 0) +
#   facet_wrap(~ term, ncol = 5) +
#   labs(x = "Longitude", y = "Latitude") +
#   theme_light(base_size = 10) +
#   ggtitle("Upper CI of time slopes by each pred grid")
# 
# # Lower CI
# ggplot(swe_coast_proj) +
#   geom_raster(data = time_slopes_by_year2, aes(x = X * 1000, y = Y * 1000, fill = estimate - std.error*1.96)) +
#   geom_sf(size = 0.3) +
#   scale_fill_gradient2(midpoint = 0) +
#   facet_wrap(~ term, ncol = 5) +
#   labs(x = "Longitude", y = "Latitude") +
#   theme_light(base_size = 10) +
#   ggtitle("Lower CI of time slopes by each pred grid")
``` 

## Visualize marginal effects

### Predict condition factor using covariates in the beginning and end of time series (kind of marginal effects with several variables changing...)

```{r "marginal effects", cache=TRUE}
# All predictors:
# biomass_her_sd_sc
# biomass_spr_sc
# biomass_spr_sd_sc
# density_cod_sc
# density_cod_rec_sc
# density_cod_sd_sc
# density_fle_sc
# density_fle_rec_sc
# density_fle_sd_sc
# density_saduria_sc
# density_saduria_rec_sc
# density_saduria_sd_sc
# depth_sc
# depth_rec_sc
# depth_sd_sc
# oxy_sc
# oxy_rec_sc
# oxy_sd_sc
# temp_sc
# temp_rec_sc
# temp_sd_sc

# Prepare prediction data frame
# nd <- data.frame(depth_sc = c(0, 1))
# 
# fix <- c(0, 0)
# 
# pred_grid_ave_1993 <- pred_grid %>%
#   filter(year == 1993) %>% 
#   dplyr::select(ends_with("sc")) %>% 
#   summarise_all(mean) %>% 
#   mutate(year = 1993L) %>% 
#   ungroup()
# 
# pred_grid_ave_2019 <- pred_grid %>%
#   filter(year == 2019) %>% 
#   dplyr::select(ends_with("sc")) %>% 
#   summarise_all(mean) %>% 
#   mutate(year = 2019L) %>% 
#   ungroup()
# 
# pred_grid_ave_covars <- bind_rows(pred_grid_ave_1993, pred_grid_ave_2019) %>% as.data.frame()
# 
# nd <- nd %>%
#   mutate(ln_length_cm = c(0, 0),
#          year = c(1993L, 2019L),
#          biomass_her_sc = pred_grid_ave_covars$biomass_her_sc,
#          biomass_her_sd_sc = pred_grid_ave_covars$biomass_her_sd_sc,
#          biomass_spr_sc = pred_grid_ave_covars$biomass_spr_sc,
#          biomass_spr_sd_sc = pred_grid_ave_covars$biomass_spr_sd_sc,
#          density_cod_sc = pred_grid_ave_covars$density_cod_sc,
#          density_cod_rec_sc = pred_grid_ave_covars$density_cod_rec_sc,
#          density_cod_sd_sc = pred_grid_ave_covars$density_cod_sd_sc,
#          density_fle_sc = pred_grid_ave_covars$density_fle_sc,
#          density_fle_rec_sc = pred_grid_ave_covars$density_fle_rec_sc,
#          density_fle_sd_sc = pred_grid_ave_covars$density_fle_sd_sc,
#          density_saduria_sc = pred_grid_ave_covars$density_saduria_sc,
#          density_saduria_rec_sc = pred_grid_ave_covars$density_saduria_rec_sc,
#          density_saduria_sd_sc = pred_grid_ave_covars$density_saduria_sd_sc,
#          depth_sc = pred_grid_ave_covars$depth_sc,
#          depth_rec_sc = pred_grid_ave_covars$depth_rec_sc,
#          depth_sd_sc = pred_grid_ave_covars$depth_sd_sc,
#          oxy_sc = pred_grid_ave_covars$oxy_sc,
#          oxy_rec_sc = pred_grid_ave_covars$oxy_rec_sc,
#          oxy_sd_sc = pred_grid_ave_covars$oxy_sd_sc,
#          temp_sc = pred_grid_ave_covars$temp_sc,
#          temp_rec_sc = pred_grid_ave_covars$temp_rec_sc,
#          temp_sd_sc = pred_grid_ave_covars$temp_sd_sc,
#            )
# 
# # Predict
# ind_nd <- predict(mfull, newdata = nd, se_fit = TRUE, re_form = NA)
# 
# ggplot(ind_nd, aes(factor(year), est, ymin = est - 1.96 * est_se, ymax = est + 1.96 * est_se)) +
#   geom_errorbar(width = 0.2) + 
#   geom_point() +
#   theme(aspect.ratio = 1)
```

### Depth

```{r marginal effects of depth, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_dep <- data.frame(depth_sc = seq(min(d$depth_sc), max(d$depth_sc), length.out = 100))

nd_dep <- nd_dep %>%
  mutate(year = 2003L,
         ln_length_cm = 0,
         biomass_her_sc = 0,
         biomass_her_sd_sc = 0,
         biomass_spr_sc = 0,
         biomass_spr_sd_sc = 0,
         density_cod_sc = 0,
         density_cod_rec_sc = 0,
         density_fle_sc = 0,
         density_fle_rec_sc = 0,
         density_saduria_sc = 0,
         density_saduria_rec_sc = 0,
         #depth_sc = 0,
         depth_rec_sc = 0,
         oxy_sc = 0,
         oxy_rec_sc = 0,
         temp_sc = 0,
         temp_rec_sc = 0
         )

# Predict from full model (AIC-selected)
p_margin_dep <- predict(mfull, newdata = nd_dep, se_fit = TRUE, re_form = NA)

mar_dep <- ggplot(p_margin_dep, aes(depth_sc, est, ymin = est - 1.96 * est_se, ymax = est + 1.96 * est_se)) +
  geom_ribbon(alpha = 0.4) +
  geom_line() +
  theme(aspect.ratio = 1) + 
  labs(x = expression(Depth[haul]))
```

### Oxygen 

```{r marginal effects of oxygen, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_oxy <- data.frame(oxy_sc = seq(min(d$oxy_sc), max(d$oxy_sc), length.out = 100))

nd_oxy <- nd_oxy %>%
  mutate(year = 2003L,
         ln_length_cm = 0,
         biomass_her_sc = 0,
         biomass_her_sd_sc = 0,
         biomass_spr_sc = 0,
         biomass_spr_sd_sc = 0,
         density_cod_sc = 0,
         density_cod_rec_sc = 0,
         density_fle_sc = 0,
         density_fle_rec_sc = 0,
         density_saduria_sc = 0,
         density_saduria_rec_sc = 0,
         depth_sc = 0,
         depth_rec_sc = 0,
         #oxy_sc = 0,
         oxy_rec_sc = 0,
         temp_sc = 0,
         temp_rec_sc = 0
         )

# Predict from full model (AIC-selected)
p_margin_oxy <- predict(mfull, newdata = nd_oxy, se_fit = TRUE, re_form = NA)

mar_oxy <- ggplot(p_margin_oxy, aes(oxy_sc, est, ymin = est - 1.96 * est_se, ymax = est + 1.96 * est_se)) +
  geom_ribbon(alpha = 0.4) +
  geom_line() +
  theme(aspect.ratio = 1) + 
  labs(x = expression(Oxygen[haul]))
```

### Temperature

```{r marginal effects of temperature, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_tem <- data.frame(temp_sc = seq(min(d$temp_sc), max(d$temp_sc), length.out = 100))

nd_tem <- nd_tem %>%
  mutate(year = 2003L,
         ln_length_cm = 0,
         biomass_her_sc = 0,
         biomass_her_sd_sc = 0,
         biomass_spr_sc = 0,
         biomass_spr_sd_sc = 0,
         density_cod_sc = 0,
         density_cod_rec_sc = 0,
         density_fle_sc = 0,
         density_fle_rec_sc = 0,
         density_saduria_sc = 0,
         density_saduria_rec_sc = 0,
         depth_sc = 0,
         depth_rec_sc = 0,
         oxy_sc = 0,
         oxy_rec_sc = 0,
         #temp_sc = 0,
         temp_rec_sc = 0
         )

# Predict from full model (AIC-selected)
p_margin_tem <- predict(mfull, newdata = nd_tem, se_fit = TRUE, re_form = NA)

mar_temp <- ggplot(p_margin_tem, aes(temp_sc, est, ymin = est - 1.96 * est_se, ymax = est + 1.96 * est_se)) +
  geom_ribbon(alpha = 0.4) +
  geom_line() +
  theme(aspect.ratio = 1) + 
  labs(x = expression(Temperature[haul]))
```

### Cod

```{r marginal effects of cod, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_cod <- data.frame(density_cod_sc = seq(min(d$density_cod_sc), max(d$density_cod_sc), length.out = 100))

nd_cod <- nd_cod %>%
  mutate(year = 2003L,
         ln_length_cm = 0,
         biomass_her_sc = 0,
         biomass_her_sd_sc = 0,
         biomass_spr_sc = 0,
         biomass_spr_sd_sc = 0,
         #density_cod_sc = 0,
         density_cod_rec_sc = 0,
         density_fle_sc = 0,
         density_fle_rec_sc = 0,
         density_saduria_sc = 0,
         density_saduria_rec_sc = 0,
         depth_sc = 0,
         depth_rec_sc = 0,
         oxy_sc = 0,
         oxy_rec_sc = 0,
         temp_sc = 0,
         temp_rec_sc = 0
         )

# Predict from full model (AIC-selected)
p_margin_cod <- predict(mfull, newdata = nd_cod, se_fit = TRUE, re_form = NA)

mar_cod <- ggplot(p_margin_cod, aes(density_cod_sc, est, ymin = est - 1.96 * est_se, ymax = est + 1.96 * est_se)) +
  geom_ribbon(alpha = 0.4) +
  geom_line() +
  theme(aspect.ratio = 1) + 
  labs(x = expression(Cod[haul]))
```

### Sprat

```{r marginal effects of sprat, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_spr <- data.frame(biomass_spr_sd_sc = seq(min(d$biomass_spr_sd_sc), max(d$biomass_spr_sd_sc), length.out = 100))

nd_spr <- nd_spr %>%
  mutate(year = 2003L,
         ln_length_cm = 0,
         biomass_her_sc = 0,
         biomass_her_sd_sc = 0,
         biomass_spr_sc = 0,
         #biomass_spr_sd_sc = 0,
         density_cod_sc = 0,
         density_cod_rec_sc = 0,
         density_fle_sc = 0,
         density_fle_rec_sc = 0,
         density_saduria_sc = 0,
         density_saduria_rec_sc = 0,
         depth_sc = 0,
         depth_rec_sc = 0,
         oxy_sc = 0,
         oxy_rec_sc = 0,
         temp_sc = 0,
         temp_rec_sc = 0
         )

# Predict from full model (AIC-selected)
p_margin_spr <- predict(mfull, newdata = nd_spr, se_fit = TRUE, re_form = NA)

mar_spr <- ggplot(p_margin_spr, aes(biomass_spr_sd_sc, est, ymin = est - 1.96 * est_se, ymax = est + 1.96 * est_se)) +
  geom_ribbon(alpha = 0.4) +
  geom_line() +
  #coord_cartesian(ylim = c(-4.65, -4.4)) + 
  theme(aspect.ratio = 1) + 
  labs(x = expression(Sprat[sub]))
```

### Saduria

```{r marginal effects of saduria, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_sad <- data.frame(density_saduria_sc = seq(min(d$density_saduria_sc), max(d$density_saduria_sc), length.out = 100))

nd_sad <- nd_sad %>%
  mutate(year = 2003L,
         ln_length_cm = 0,
         biomass_her_sc = 0,
         biomass_her_sd_sc = 0,
         biomass_spr_sc = 0,
         biomass_spr_sd_sc = 0,
         density_cod_sc = 0,
         density_cod_rec_sc = 0,
         density_fle_sc = 0,
         density_fle_rec_sc = 0,
         #density_saduria_sc = 0,
         density_saduria_rec_sc = 0,
         depth_sc = 0,
         depth_rec_sc = 0,
         oxy_sc = 0,
         oxy_rec_sc = 0,
         temp_sc = 0,
         temp_rec_sc = 0
         )

# Predict from full model (AIC-selected)
p_margin_sad <- predict(mfull, newdata = nd_sad, se_fit = TRUE, re_form = NA)

mar_sad <- ggplot(p_margin_sad, aes(density_saduria_sc, est, ymin = est - 1.96 * est_se, ymax = est + 1.96 * est_se)) +
  geom_ribbon(alpha = 0.4) +
  geom_line() +
  #coord_cartesian(ylim = c(-4.75, -4.4)) + 
  theme(aspect.ratio = 1) + 
  labs(x = expression(Saduria~entomon[haul]))
```

### Plot together

```{r}
mar_dep + mar_oxy + mar_temp + mar_cod + mar_spr + mar_sad
  plot_annotation(tag_levels = 'A') 

ggsave("figures/supp/condition/marginal_effects.png", width = 6.5, height = 6.5, dpi = 600)
```

```{r}
knitr::knit_exit()
```
