---
title: "Cod cpue in relation to oxygen"
author: "Max Lindmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  df_print: paged
pdf_document: default
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.asp = 0.618,
  fig.align ='center'
)
```

# Read cpue data and fit model
```{r lib, message=FALSE}
# Load libraries, install if needed
library(tidyverse); theme_set(theme_classic())
library(readxl)
library(tidylog)
library(RCurl)
library(viridis)
library(RColorBrewer)
library(patchwork)
library(janitor)
library(icesDatras)
library(mapdata)
library(patchwork)
library(rgdal)
library(raster)
library(sf)
library(rgeos)
library(chron)
library(lattice)
library(ncdf4)
library(sdmTMB) # remotes::install_github("pbs-assess/sdmTMB")
library(marmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(mapplots)
library(qwraps2) # To load entire cache in interactive r session, do: qwraps2::lazyload_cache_dir(path = "R/analysis/cpue_cod_cache/html")
```

Read data:

```{r read and process data, message=FALSE, warning=FALSE}
d <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/mdat_cpue.csv")

# Calculate standardized variables
d <- d %>% 
  mutate(ln_length_cm = log(length_cm),
         ln_weight_g = log(weight_g),
         oxy_st = oxy,
         oxy_rec_st = oxy_rec,
         temp_st = temp,
         temp_rec_st = temp_rec,
         abun_her_st = abun_her,
         abun_her_sd_st = abun_her_sd,
         abun_spr_st = abun_spr,
         abun_spr_sd_st = abun_spr_sd,
         cpue_cod_st = cpue_cod,
         cpue_cod_rec_st = cpue_cod_rec,
         cpue_fle_st = cpue_fle,
         cpue_fle_rec_st = cpue_fle_rec,
         depth_st = depth) %>%
  mutate_at(c("oxy_st", "oxy_rec_st", "temp_st", "temp_rec_st",
              "abun_her_st", "abun_her_sd_st", "abun_spr_st", "abun_spr_sd_st",
              "cpue_cod_st", "cpue_cod_rec_st", "cpue_fle_st", "cpue_fle_rec_st",
              "depth_st"),
            ~(scale(.) %>% as.vector)) %>% 
  mutate(year = as.integer(year))

unique(is.na(d$abun_spr_st))

d %>% drop_na(abun_spr_st)
# almost 3k data points (3%) when dropping NA sprat... So we remove it?
```


Read the prediction grids:

```{r read and process prediction grid, message=FALSE, warning=FALSE}
# And now read in pred_grid2 which has oxygen values at location and time and depth:
pred_grid2 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/pred_grid2.csv")

pred_grid2 <- pred_grid2 %>%
  mutate(ln_length_cm = log(1)) %>% # For now we'll predict changes in the intercept ("condition factor")
  mutate(X = lon, Y = lat, year = as.integer(year)) %>% 
  filter(year %in% c(unique(cod$year))) %>% 
  mutate(depth_st = (depth - mean(cod$depth))/sd(cod$depth),
         oxy_st = (oxy - mean(cod$oxy))/sd(cod$oxy)) # Need to scale these to the mean and sd in the data!
```

Standardize data with respect to prediction grid:

```{r prep data for cod models, cache=TRUE}
# Standardize variables with respect to the prediction grid
pred_grid2 <- pred_grid2 %>% drop_na(depth)

cod <- cod %>%
  drop_na(depth) %>% 
  mutate(depth_st = (depth - mean(pred_grid2$depth))/sd(pred_grid2$depth),
         depth_st_sq = depth_st*depth_st)
```

Make barrier spde mesh:

```{r make barrier spde mesh, results='hide', cache=TRUE, message=FALSE}
# Crop the polygon for plotting and efficiency:
baltic_coast <- suppressWarnings(suppressMessages(
  st_crop(world,
          c(xmin = xmin, ymin = ymin, xmax = xmax, ymax = ymax))))

crs <- 4326 # https://en.wikipedia.org/wiki/EPSG_Geodetic_Parameter_Dataset#Common_EPSG_codes, WGS84

st_crs(baltic_coast) <- 4326 # 'WGS84'; necessary on some installs
baltic_coast <- st_transform(baltic_coast, crs)

# Project our survey data coordinates:
survey <- cod %>% dplyr::select(lon, lat, cpue) %>%
  st_as_sf(crs = 4326, coords = c("lon", "lat")) 

# Prepare for making the mesh
# First, we will extract the coordinates:
surv_coords <- st_coordinates(survey)

spde <- make_mesh(cod, xy_cols = c("lon", "lat"),
                  n_knots = 200, 
                  type = "kmeans", seed = 42)

# Add on the barrier mesh component:
bspde <- add_barrier_mesh(
  spde, baltic_coast, range_fraction = 0.2,
  proj_scaling = 1, plot = TRUE
)

# In the above, the grey dots are the centre of triangles that are in the
# ocean. The red crosses are centres of triangles that are over land. The
# spatial range will be assumed to be 0.2 (`range_fraction`) over land compared
# to over water.

# We can make a more advanced plot if we want:
# mesh_df_water <- bspde$mesh_sf[bspde$normal_triangles, ]
# mesh_df_land <- bspde$mesh_sf[bspde$barrier_triangles, ]

# Now, when we fit our model with the new mesh, it will automatically
# include a barrier structure in the spatial correlation:
```

Fit the models of cpue

######### RTEMOVE INTERCEPT FORM THESE MODELS1!!!!!!!!

```{r fit, results='hide', cache=TRUE, message=FALSE}
# Depth squared
m1 <- sdmTMB(cpue ~ 0 + as.factor(year) + depth_st + depth_st_sq, data = cod,
             spde = bspde, family = tweedie(link = "log"),
             ar1_fields = TRUE, include_spatial = TRUE, time = "year",
             spatial_only = FALSE, newton_steps = 1, reml = FALSE)

# Depth spline
m2 <- sdmTMB(cpue ~ 0 + as.factor(year) + s(depth_st, k = 3), data = cod,
             spde = bspde, family = tweedie(link = "log"),
             ar1_fields = TRUE, include_spatial = TRUE, time = "year",
             spatial_only = FALSE, newton_steps = 1, reml = FALSE)

# Depth squared + oxy squared
m3 <- sdmTMB(cpue ~ 0 + as.factor(year) + depth_st + depth_st_sq + oxy_st + oxy_st_sq, data = cod,
             spde = bspde, family = tweedie(link = "log"),
             ar1_fields = TRUE, include_spatial = TRUE, time = "year",
             spatial_only = FALSE, newton_steps = 1, reml = FALSE)

# Depth spline + oxy spline
m4 <- sdmTMB(cpue ~ 0 + as.factor(year) + s(depth_st, k = 3) + s(oxy_st, k = 3), data = cod,
             spde = bspde, family = tweedie(link = "log"),
             ar1_fields = TRUE, include_spatial = TRUE, time = "year",
             spatial_only = FALSE, newton_steps = 1, reml = FALSE)

#tidy(mcod, conf.int = TRUE)
cod$residualsm1 <- residuals(m1)
cod$residualsm2 <- residuals(m2)
cod$residualsm3 <- residuals(m3)
cod$residualsm4 <- residuals(m4)
qqnorm(cod$residualsm1); abline(a = 0, b = 1)
qqnorm(cod$residualsm2); abline(a = 0, b = 1)
qqnorm(cod$residualsm3); abline(a = 0, b = 1)
qqnorm(cod$residualsm4); abline(a = 0, b = 1)

AIC(m1, m2, m3, m4) %>% arrange(AIC)
```

Check the AR1 parameter (`rho` is `ar_phi` on the -1 to 1 scale):

```{r check AR1 estimate}
tidy(m4, effects = "ran_pars", conf.int = TRUE) %>% filter(term == "rho")
```

Predict and extract CPUE-weighted mean oxygen and depth per prediction grid

```{r predict}
pred_grid2 <- pred_grid2 %>%
  mutate(X = lon, Y = lat, year = as.integer(year)) %>% 
  filter(year %in% c(unique(cod$year))) %>% 
  mutate(depth_st = (depth - mean(cod$depth))/sd(cod$depth),
         oxy_st = (oxy - mean(cod$oxy))/sd(cod$oxy),
         depth_st_sq = depth_st*depth_st) %>% # Need to scale these to the mean and sd in the data!
  drop_na(oxy_st)

predcod <- predict(m4, newdata = pred_grid2)
```

Plot on map

```{r plot}
predcod <- predcod %>% mutate(est2 = ifelse(depth > 120, NA, est))

# predcod %>% mutate(est2_resp = exp(est2)) %>% arrange(exp(est2_resp)) %>% dplyr::select(est2_resp) %>% head(20)
# predcod %>% mutate(est2_resp = exp(est2)) %>% arrange(desc(est2_resp)) %>% dplyr::select(est2_resp) %>% head(20)

# Oxygen
predcod %>% 
  #filter(year %in% c("1993", "2018")) %>% 
  filter(exp(est2) < 2000) %>% 
  filter(exp(est2) > 0.01) %>% # remove some extremes to better see trends in space and time
  ggplot(., aes(X, Y, fill = exp(est2))) +
  geom_raster() +
  facet_wrap(~year, ncol = 5) +
  scale_fill_viridis(option = "magma", 
                     name = "cpue") + 
  geom_sf(data = world, inherit.aes = F, size = 0.2) +
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  ggtitle("Prediction (random + fixed)") +
  labs(x = "lon", y = "lat")
```

We can also plot the annual index

```{r annual condition factor, message=FALSE, cache=TRUE}
# From these models, predict annual condition factor
# Grabbing the number of cells to help with calculating the average
ncells <- filter(pred_grid2, year == max(pred_grid2$year)) %>% nrow()

# Use the `area` argument here to turn the total into an average by giving it one over the number of cells
pred_grid2 <- pred_grid2 %>% filter(depth < 120)

preds_mcod <- predict(mcod, newdata = pred_grid2, return_tmb_object = TRUE, area = 1/ncells)

# Make a little helper function... bias correction shouldn't do anything here because of the identity link
get_average_index <- function(obj, level = 0.95, ...)  {
  sdmTMB:::get_generic(obj, value_name = "link_total",
    bias_correct = FALSE, level = level, trans = I, ...)
}

avg_mcod <- get_average_index(preds_mcod)

avg_mcod <- avg_mcod %>% mutate(est_response = exp(est),
                                lwr_response = exp(lwr),
                                upr_response = exp(upr))

avg_mcod %>%
  ggplot(., aes(year, est_response)) +
  ylab("cpue") +
  geom_line() + 
  geom_ribbon(aes(ymin = lwr_response, ymax = upr_response), alpha = 0.4) +
  theme(axis.text.x = element_text(angle = 30),
        legend.position = c(0.8, 0.8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(x = "Year") +
  NULL
```

Calculate cpue-weighted mean oxygen-at-location

```{r cpue-weighted mean oxygen plot}
predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  #filter(exp(est2) < 1000) %>% # remove some extremes to better see trends in space and time 
  #filter(exp(est2) > 0.05) %>% # remove some extremes to better see trends in space and time
  summarise(oxygen_wm = weighted.mean(oxy, exp(est2))) %>% 
  ggplot(., aes(year, oxygen_wm)) +
  geom_point(size = 3) +
  stat_smooth(method = "lm") +
  theme(axis.text.x = element_text(angle = 30),
        legend.position = c(0.8, 0.8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  ggtitle("CPUE weighted average oxygen at location") 
```

Calculate cpue-weighted mean depth-at-location
```{r cpue-weighted mean depth plot}
predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  #filter(exp(est2) < 1000) %>% # remove some extremes to better see trends in space and time 
  #filter(exp(est2) > 0.05) %>% # remove some extremes to better see trends in space and time
  summarise(depth_wm = weighted.mean(depth, exp(est2))) %>% 
  ggplot(., aes(year, depth_wm)) +
  geom_point(size = 3) +
  stat_smooth(method = "lm") +
  theme(axis.text.x = element_text(angle = 30),
        legend.position = c(0.8, 0.8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  ggtitle("CPUE weighted average depth at location") 
```

Marginal effect of depth

```{r marginal effects depth, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_dep <- data.frame(depth = seq(min(cod$depth), max(cod$depth), length.out = 100))
nd_dep <- nd %>%
  mutate(year = 2003L,
         depth_st = (depth - mean(depth))/sd(depth),
         depth_st_sq = depth_st*depth_st,
         oxy_st = 0,
         oxy_st_sq = 0)

# Predict 
p_margin_dep <- predict(m4, newdata = nd_dep, se_fit = TRUE, re_form = NA)

ggplot(p_margin_dep, aes(depth, exp(est),
  ymin = exp(est) - 1.96 * exp(est_se), ymax = exp(est) + 1.96 * exp(est_se))) +
  geom_line() + 
  geom_ribbon(alpha = 0.4)
```

Marginal effect of oxygen

```{r marginal effects oxygen, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_oxy <- data.frame(oxy = seq(min(cod$oxy), max(cod$oxy), length.out = 100))
nd_oxy <- nd_oxy %>%
  mutate(year = 2003L,
         depth_st = 0,
         depth_st_sq = 0,
         oxy_st = (oxy - mean(oxy))/sd(oxy),
         oxy_st_sq = oxy_st*oxy_st)

# Predict 
p_margin_oxy <- predict(m4, newdata = nd_oxy, se_fit = TRUE, re_form = NA)

ggplot(p_margin_oxy, aes(oxy, exp(est),
  ymin = exp(est) - 1.96 * exp(est_se), ymax = exp(est) + 1.96 * exp(est_se))) +
  geom_line() + 
  geom_ribbon(alpha = 0.4)
```

