---
title: "Cod cpue in relation to oxygen"
author: "Max Lindmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  df_print: paged
pdf_document: default
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.asp = 0.618,
  fig.align ='center'
)
```

# Read cpue data and fit model
```{r lib, message=FALSE}
# Load libraries, install if needed
library(tidyverse); theme_set(theme_classic())
library(readxl)
library(tidylog)
library(RCurl)
library(viridis)
library(RColorBrewer)
library(patchwork)
library(janitor)
library(icesDatras)
library(mapdata)
library(patchwork)
library(rgdal)
library(raster)
library(sf)
library(rgeos)
library(chron)
library(lattice)
library(ncdf4)
library(sdmTMB) # remotes::install_github("pbs-assess/sdmTMB")
library(marmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(mapplots)
library(qwraps2) # To load entire cache in interactive r session, do: qwraps2::lazyload_cache_dir(path = "R/analysis/cpue_model_cache/html")

# For adding maps to plots
world <- ne_countries(scale = "medium", returnclass = "sf")

# Specify map ranges
ymin = 54; ymax = 58; xmin = 9.5; xmax = 22
```

Read data:

```{r read and process data, message=FALSE, warning=FALSE}
d <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/mdat_cpue.csv")

# Calculate standardized variables
d <- d %>% 
  mutate(oxy_sc = oxy,
         temp_sc = temp,
         depth_sc = depth,
         ) %>%
  mutate_at(c("oxy_sc", "temp_sc", "depth_sc"),
            ~(scale(.) %>% as.vector)) %>% 
  mutate(year = as.integer(year)) %>% 
  drop_na(oxy, depth, temp)

ggplot(d, aes(lon, lat)) + geom_point(size = 0.3) + facet_wrap(~year)
```

Read the prediction grids:

```{r read and process prediction grid, message=FALSE, warning=FALSE}
# And now read in pred_grid2 which has oxygen values at location and time and depth:
pred_grid2 <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/pred_grid2.csv")


# Standardize data with respect to prediction grid:
pred_grid2 <- pred_grid2 %>%
  mutate(ln_length_cm = log(1)) %>% # For now we'll predict changes in the intercept ("condition factor")
  mutate(X = lon, Y = lat, year = as.integer(year)) %>% 
  filter(year %in% c(unique(d$year))) %>% 
  mutate(depth_sc = (depth - mean(d$depth))/sd(d$depth),
         temp_sc = (temp - mean(d$temp))/sd(d$temp),
         oxy_sc = (oxy - mean(d$oxy))/sd(d$oxy)) # Need to scale these to the mean and sd in the data!
```

Make barrier spde mesh:

```{r make barrier spde mesh, results='hide', cache=TRUE, message=FALSE}
# Crop the polygon for plotting and efficiency:
baltic_coast <- suppressWarnings(suppressMessages(
  st_crop(world,
          c(xmin = xmin, ymin = ymin, xmax = xmax, ymax = ymax))))

crs <- 4326 # https://en.wikipedia.org/wiki/EPSG_Geodetic_Parameter_Dataset#Common_EPSG_codes, WGS84

st_crs(baltic_coast) <- 4326 # 'WGS84'; necessary on some installs
baltic_coast <- st_transform(baltic_coast, crs)

# Project our survey data coordinates:
survey <- d %>% dplyr::select(lon, lat, cpue) %>%
  st_as_sf(crs = 4326, coords = c("lon", "lat")) 

# Prepare for making the mesh
# First, we will extract the coordinates:
surv_coords <- st_coordinates(survey)

spde <- make_mesh(d, xy_cols = c("lon", "lat"),
                  n_knots = 200, 
                  type = "kmeans", seed = 42)

plot(spde)

# Add on the barrier mesh component:
bspde <- add_barrier_mesh(
  spde, baltic_coast, range_fraction = 0.2,
  proj_scaling = 1, plot = TRUE
)

# In the above, the grey dots are the centre of triangles that are in the
# ocean. The red crosses are centres of triangles that are over land. The
# spatial range will be assumed to be 0.2 (`range_fraction`) over land compared
# to over water.

# We can make a more advanced plot if we want:
# mesh_df_water <- bspde$mesh_sf[bspde$normal_triangles, ]
# mesh_df_land <- bspde$mesh_sf[bspde$barrier_triangles, ]

# Now, when we fit our model with the new mesh, it will automatically
# include a barrier structure in the spatial correlation:
```

Fit the models of cpue:

```{r fit, results='hide', cache=TRUE, message=FALSE}
# Depth spline + oxy spline
# Takes about 30 minutes
m <- sdmTMB(cpue ~ 0 + as.factor(year) + s(depth_sc, k = 3) + s(oxy_sc, k = 3) + s(temp_sc, k = 3),
            data = d, spde = bspde, family = tweedie(link = "log"),
            fields = "AR1", include_spatial = TRUE, time = "year",
            spatial_only = FALSE, reml = FALSE,
            control = sdmTMBcontrol(newton_steps = 1))

tidy(m, conf.int = TRUE)
d$residualsm <- residuals(m)
qqnorm(d$residualsm); abline(a = 0, b = 1)
```

Check the AR1 parameter (`rho` is `ar_phi` on the -1 to 1 scale):

```{r check AR1 estimate}
m <- sdmTMB:::update_model(m) # Update the model formula instead...
  
tidy(m, effects = "ran_pars", conf.int = TRUE) %>% filter(term == "rho")
```

Predict and extract CPUE-weighted mean oxygen and depth per prediction grid

```{r predict, cache=TRUE}
predcod <- predict(m, newdata = pred_grid2)

predcod <- predcod %>% mutate(est2 = ifelse(depth > 120, NA, est))
```

Plot on map

```{r plot}
# Plot predicted CPUE
predcod %>% 
  #filter(year %in% c("1993", "2018")) %>% 
  filter(exp(est2) < 2000) %>% 
  filter(exp(est2) > 0.01) %>% # remove some extremes to better see trends in space and time
  ggplot(., aes(X, Y, fill = exp(est2))) +
  geom_raster() +
  facet_wrap(~year, ncol = 5) +
  scale_fill_viridis(option = "magma", 
                     name = "cpue") + 
  geom_sf(data = world, inherit.aes = F, size = 0.2) +
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  ggtitle("Prediction (random + fixed)") +
  labs(x = "lon", y = "lat")
```

We can also plot the annual index

```{r cpue index, message=FALSE, cache=TRUE}
# From these models, predict annual cpue
# Grabbing the number of cells to help with calculating the average
pred_grid2 <- pred_grid2 %>% filter(depth < 120)

unique(is.na(pred_grid2))
 
pred_grid2 <- drop_na(pred_grid2)

ncells <- filter(pred_grid2, year == max(pred_grid2$year)) %>% nrow()

preds_mcod <- predict(m, newdata = pred_grid2, return_tmb_object = TRUE)

ind <- get_index(preds_mcod)

# The total area is 4053 cells * 36.4 km^2 of a cell
ind <- ind %>% mutate(est_kg_h = est/(ncells*36.4),
                      lwr_kg_h = lwr/(ncells*36.4),
                      upr_kg_h = upr/(ncells*36.4))
 
ind %>%
  ggplot(., aes(year, est_kg_h)) +
  ylab("Average CPUE [kg/hour]") +
  geom_line() + 
  geom_ribbon(aes(ymin = lwr_kg_h, ymax = upr_kg_h), alpha = 0.4) +
  theme(axis.text.x = element_text(angle = 30),
        legend.position = c(0.8, 0.8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(x = "Year") +
  NULL
```

Calculate cpue-weighted mean oxygen-at-location

```{r cpue-weighted mean oxygen plot}
aa <- predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  #filter(exp(est2) < 1000) %>% # remove some extremes to better see trends in space and time 
  #filter(exp(est2) > 0.05) %>% # remove some extremes to better see trends in space and time
  summarise(oxygen_wm = weighted.mean(oxy, exp(est2))) %>% 
  ggplot(., aes(year, oxygen_wm)) +
  geom_point(size = 3) +
  stat_smooth(method = "lm") +
  theme(axis.text.x = element_text(angle = 30),
        legend.position = c(0.8, 0.8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  ggtitle("CPUE weighted average oxygen at location") 

# Only SD 25
bb <- predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  filter(lon > 15 & lon < 18 & lat < 56.5) %>% 
  summarise(oxygen_wm = weighted.mean(oxy, exp(est2))) %>% 
  ggplot(., aes(year, oxygen_wm)) +
  geom_point(size = 3) +
  stat_smooth(method = "lm") +
  theme(axis.text.x = element_text(angle = 30),
        legend.position = c(0.8, 0.8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  ggtitle("SD25: CPUE weighted average oxygen at location") 

aa / bb
```

Calculate cpue-weighted mean depth-at-location

```{r cpue-weighted mean depth plot}
predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  #filter(exp(est2) < 1000) %>% # remove some extremes to better see trends in space and time 
  #filter(exp(est2) > 0.05) %>% # remove some extremes to better see trends in space and time
  summarise(depth_wm = weighted.mean(depth, exp(est2))) %>% 
  ggplot(., aes(year, depth_wm)) +
  geom_point(size = 3) +
  stat_smooth(method = "lm") +
  theme(axis.text.x = element_text(angle = 30),
        legend.position = c(0.8, 0.8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  ggtitle("CPUE weighted average depth at location") 
```

Calculate weighted mean depth

```{r stats on mean depth}
dep <- predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  #filter(exp(est2) < 1000) %>% # remove some extremes to better see trends in space and time 
  #filter(exp(est2) > 0.05) %>% # remove some extremes to better see trends in space and time
  summarise(depth_wm = weighted.mean(depth, exp(est2))) %>% 
  mutate(year_ct = year - 1993)

summary(lm(dep$depth_wm ~ dep$year_ct))

46.06615 + 0.18151*(2019-1993)

# Now plot oxygen vs depth
# Plot oxygen vs depth
ggplot(pred_grid2, aes(depth, oxy)) + geom_point() + stat_smooth(method = "lm")
ggplot(pred_grid2, aes(depth, oxy)) +
  geom_point() +
  stat_smooth(method = "lm") +
  coord_cartesian(xlim = c(46, 51), ylim = c(5.4, 5.7))

```

Calculate cpue-weighted mean temperature-at-location

```{r cpue-weighted mean temperature plot}
predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  #filter(exp(est2) < 1000) %>% # remove some extremes to better see trends in space and time 
  #filter(exp(est2) > 0.05) %>% # remove some extremes to better see trends in space and time
  summarise(temp_wm = weighted.mean(temp, exp(est2))) %>% 
  ggplot(., aes(year, temp_wm)) +
  geom_point(size = 3) +
  stat_smooth(method = "gam", formula = y ~ s(x, k = 3)) +
  theme(axis.text.x = element_text(angle = 30),
        legend.position = c(0.8, 0.8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  ggtitle("CPUE weighted average temp at location") 
```

Combined plot:

```{r combined plot, message=FALSE}
cpue_index <- ind %>%
  ggplot(., aes(year, est_kg_h)) +
  ylab("Mean CPUE [kg/hour]") +
  geom_line() + 
  geom_ribbon(aes(ymin = lwr_kg_h, ymax = upr_kg_h), alpha = 0.4) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(x = "Year") +
  theme_classic(base_size = 13) +
  theme(axis.text.x = element_text(angle = 20, size = 7))

plot_dep <- predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  summarise(depth_wm = weighted.mean(depth, exp(est2))) %>%
  #summarise(depth_wm = weighted.mean(depth, est2)) %>% 
  ggplot(., aes(year, depth_wm)) +
  geom_point(size = 1.8) +
  labs(y = "Depth [m]", x = "Year") + 
  stat_smooth(method = "gam", color = "grey20", formula = y ~ s(x, k = 3)) +
  #stat_smooth(method = "gam", color = "grey20") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_classic(base_size = 13) +
  theme(axis.text.x = element_text(angle = 20, size = 7))

# For these two plots we want to calculate the average in the prediction grid and contrast
# that to the biomass-weighted
annual_oxy_temp <- pred_grid2 %>% 
  drop_na(oxy, temp) %>% 
  mutate(Year = year) %>% 
  group_by(Year) %>% 
  summarize(mean_temp = mean(temp),
            mean_oxy = mean(oxy)) %>% 
  ungroup()

# ggplot(annual_oxy_temp, aes(Year, mean_oxy)) + geom_point() + stat_smooth(method = "lm")
  
pal <- RColorBrewer::brewer.pal(name = "Dark2", n = 5)

plot_oxy <- predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  summarise(oxy_wm = weighted.mean(oxy, exp(est2))) %>% 
  #summarise(oxy_wm = weighted.mean(oxy, est2)) %>% 
  ggplot(.) +
  geom_point(aes(year, oxy_wm, color = "red"), size = 1.8, show.legend = TRUE) +
  geom_point(data = annual_oxy_temp, aes(Year, mean_oxy, color = "blue"), size = 1.8, show.legend = TRUE) +
  #stat_smooth(aes(year, oxy_wm), method = "lm", fill = pal[2], alpha = 0.2, color = pal[2]) + # lm
  stat_smooth(aes(year, oxy_wm), method = "gam", formula = y ~ s(x, k = 10), fill = pal[2], alpha = 0.2, color = pal[2]) + # gam
  #stat_smooth(data = annual_oxy_temp, aes(Year, mean_oxy), method = "lm", fill = pal[3], alpha = 0.2, color = pal[3]) +
  stat_smooth(data = annual_oxy_temp, aes(Year, mean_oxy), method = "gam",
              formula = y ~ s(x, k = 3), fill = pal[3], alpha = 0.2, color = pal[3]) + # gam
  scale_color_manual(values = c(pal[3], pal[2]),
                     name = c(""),
                     labels = c("Environment", "CPUE-weighted")) +
  labs(y = expression(paste("Mean O" [2], " [ml/L]", sep = "")), x = "Year") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_classic(base_size = 13) +
  guides(colour = guide_legend(override.aes = list(size = 1)))+
  theme(legend.position = c(0.8, 0.95),
        legend.text = element_text(size = 7),
        axis.text.x = element_text(angle = 20, size = 7),
        legend.background = element_rect(fill = NA),
        legend.margin = margin(0, 0, 0, 0),
        legend.spacing.x = unit(0, "pt"))

plot_temp <- predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  summarise(temp_wm = weighted.mean(temp, exp(est2))) %>%
  #summarise(temp_wm = weighted.mean(temp, est2)) %>% 
  ggplot(.) +
  geom_point(aes(year, temp_wm, color = "red"), size = 1.8, show.legend = TRUE) +
  geom_point(data = annual_oxy_temp, aes(Year, mean_temp, color = "blue"), size = 1.8, show.legend = TRUE) +
  #stat_smooth(aes(year, temp_wm), method = "lm", fill = pal[2], alpha = 0.2, color = pal[2]) + # lm
  stat_smooth(aes(year, temp_wm), method = "gam", formula = y ~ s(x, k = 3),
              fill = pal[2], alpha = 0.2, color = pal[2]) + # gam
  #stat_smooth(data = annual_oxy_temp, aes(Year, mean_temp), method = "lm", fill = pal[3], alpha = 0.2, color = pal[3]) +
  stat_smooth(data = annual_oxy_temp, aes(Year, mean_temp), method = "gam", formula = y ~ s(x, k = 3),
              fill = pal[3], alpha = 0.2, color = pal[3]) +
  scale_color_manual(values = c(pal[3], pal[2]),
                     name = c(""),
                     labels = c("Environment", "CPUE-weighted")) +
  labs(y = "Temperature [°C]", x = "Year") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_classic(base_size = 13) +
  guides(colour = guide_legend(override.aes = list(size = 1)))+
  theme(legend.position = c(0.25, 0.95),
        legend.text = element_text(size = 7),
        axis.text.x = element_text(angle = 20, size = 7),
        legend.background = element_rect(fill = NA),
        legend.margin = margin(0, 0, 0, 0),
        legend.spacing.x = unit(0, "pt"))

(cpue_index + plot_dep) / (plot_oxy + plot_temp) +
  plot_annotation(tag_levels = 'A')

ggsave("figures/analysis/cpue.png", width = 6.5, height = 6.5, dpi = 600)
```

Calculate slopes of average oxygen etc.

```{r simple stats}
# Do some simple stats...
annual_oxy_temp$source <- "env"

weigh <- predcod %>% 
  drop_na(est2) %>% 
  group_by(year) %>% 
  summarise(mean_temp = weighted.mean(temp, exp(est2)),
            mean_oxy = weighted.mean(oxy, exp(est2))) %>% 
  mutate(source = "cpue_model") %>% 
  rename("Year" = "year")
  
dd <- bind_rows(annual_oxy_temp, weigh)

ggplot(dd, aes(Year, mean_temp, color = source)) +
  geom_point() + 
  stat_smooth(method = "lm")

ggplot(dd, aes(Year, mean_oxy, color = source)) +
  geom_point() + 
  stat_smooth(method = "lm")

summary(lm(mean_temp ~ Year*source, data = dd))
summary(lm(mean_oxy ~ Year*source, data = dd))

summary(lm(mean_oxy ~ Year, data = subset(dd, source == "env")))
summary(lm(mean_oxy ~ Year, data = subset(dd, source == "cpue_model")))
```

Now calculate the average cod density by sub-divisions and year, and compare that to the pelagic density

```{r calculate cod-pelagics ratio}
# Read the condition data which has the pelagics abundances
cond <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/cod_condition/master/data/for_analysis/mdat_cond.csv")

# Add in ICES rectangle information
cond$ices_rect <- ices.rect2(lon = cond$lon, lat = cond$lat)
predcod$ices_rect <- ices.rect2(lon = predcod$lon, lat = predcod$lat) # predcod is predicted cod cpue across grid

# Add in ICES sub-division information
cond <- cond %>% 
  mutate(SubDiv = NA) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("37G0", "37G1",
                                          "38G0", "38G1", 
                                          "39F9", "39G0", "39G1",
                                          "40F9", "40G0", "40G1"), "22", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect == "40G2", "23", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("37G2", "37G3", "37G4",
                                          "38G1", "38G2", "38G3", "38G4", 
                                          "39G1", "39G2", "39G3", "39G4",
                                          "40G1"), "24", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("40G4",
                                          "37G5", "37G6", "37G7",
                                          "38G5", "38G6", "38G7",
                                          "39G5", "39G6", "39G7",
                                          "40G5", "40G6", "40G7",
                                          "41G5", "41G6", "41G7"), "25", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("37G8", "37G9", "37H0",
                                          "38G8", "38G9", "38H0",
                                          "39G8", "39G9", "39H0",
                                          "40G8", "40G9", "40H0",
                                          "41G8", "41G9", "41H0"), "26", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("42G6", "42G7",
                                          "43G6", "43G7",
                                          "44G6", "44G7", "44G8"), "27", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("42G8", "42G9", "42H0", "42H1", "42H2",
                                          "43G8", "43G9", "43H0", "43H1", "43H2",
                                          "44G8", "44G9", "44H0", "44H1", "44H2"), "28", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 58.5 & lon > 19, "29", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 57 & lat < 58.5 & lon > 19 & lon < 22, "28", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 57 & lat < 58.5 & lon > 18 & lon < 22, "28", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 57 & lat < 60 & lon > 16 & lon < 18, "27", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 55.5 & lat < 56.5 & lon > 14 & lon < 16, "25", SubDiv))

# Now for predicted CPUE
predcod <- predcod %>% 
  mutate(SubDiv = NA) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("37G0", "37G1",
                                          "38G0", "38G1", 
                                          "39F9", "39G0", "39G1",
                                          "40F9", "40G0", "40G1"), "22", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect == "40G2", "23", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("37G2", "37G3", "37G4",
                                          "38G1", "38G2", "38G3", "38G4", 
                                          "39G1", "39G2", "39G3", "39G4",
                                          "40G1"), "24", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("40G4",
                                          "37G5", "37G6", "37G7",
                                          "38G5", "38G6", "38G7",
                                          "39G5", "39G6", "39G7",
                                          "40G5", "40G6", "40G7",
                                          "41G5", "41G6", "41G7"), "25", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("37G8", "37G9", "37H0",
                                          "38G8", "38G9", "38H0",
                                          "39G8", "39G9", "39H0",
                                          "40G8", "40G9", "40H0",
                                          "41G8", "41G9", "41H0"), "26", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("42G6", "42G7",
                                          "43G6", "43G7",
                                          "44G6", "44G7", "44G8"), "27", SubDiv)) %>% 
  mutate(SubDiv = ifelse(ices_rect %in% c("42G8", "42G9", "42H0", "42H1", "42H2",
                                          "43G8", "43G9", "43H0", "43H1", "43H2",
                                          "44G8", "44G9", "44H0", "44H1", "44H2"), "28", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 58.5 & lon > 19, "29", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 57 & lat < 58.5 & lon > 19 & lon < 22, "28", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 57 & lat < 58.5 & lon > 18 & lon < 22, "28", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 57 & lat < 60 & lon > 16 & lon < 18, "27", SubDiv)) %>%
  mutate(SubDiv = ifelse(lat > 55.5 & lat < 56.5 & lon > 14 & lon < 16, "25", SubDiv))

# Plot and inspect
ggplot(cond, aes(x = lon, y = lat, color = SubDiv)) +
  geom_point() +
  geom_sf(data = world, inherit.aes = F, size = 0.2) +
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  NULL

ggplot(predcod, aes(x = lon, y = lat, color = SubDiv)) +
  geom_point() +
  geom_sf(data = world, inherit.aes = F, size = 0.2) +
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  NULL

# Now calculate average sprat and herring density by sub-div
pelagics <- cond %>%
  drop_na(SubDiv, abun_spr_sd, abun_her_sd) %>% 
  group_by(year, SubDiv) %>%
  summarise(ave_abun_spr = mean(abun_spr_sd),
            ave_abun_her = mean(abun_her_sd)) %>%
  ungroup() %>% 
  dplyr::select(year, SubDiv, ave_abun_spr, ave_abun_her) %>% 
  mutate(id = paste(year, SubDiv, sep = "_"))

# And average CPUE
cod <- predcod %>%
  drop_na(SubDiv, est) %>% 
  group_by(year, SubDiv) %>%
  summarise(ave_cpue_cod = mean(est)) %>%
  ungroup() %>% 
  mutate(id = paste(year, SubDiv, sep = "_")) %>% 
  dplyr::select(id, ave_cpue_cod)
  
dat <- left_join(pelagics, cod)

head(dat)

dat <- dat %>% 
  mutate(ave_abun_spr_sc = ave_abun_spr,
         ave_abun_her_sc = ave_abun_her,
         ave_cpue_cod_sc = ave_cpue_cod) %>%
  mutate_at(c("ave_abun_spr_sc", "ave_abun_her_sc", "ave_cpue_cod_sc"), ~(scale(.) %>% as.vector))

dat$cod_spr_ratio <- dat$ave_cpue_cod / dat$ave_abun_spr
dat$cod_her_ratio <- dat$ave_cpue_cod / dat$ave_abun_her

dat$cod_spr_ratio_sc <- dat$ave_cpue_cod_sc / dat$ave_abun_spr_sc
dat$cod_her_ratio_sc <- dat$ave_cpue_cod_sc / dat$ave_abun_her_sc

# Non-scaled variables
p1 <- ggplot(dat, aes(year, log(cod_her_ratio), color = SubDiv)) +
  geom_point() +
  stat_smooth() +
  scale_color_brewer(palette = "Dark2") +
  ggtitle("Herring") +
  theme(aspect.ratio = 3/4)

# Sprat ratio
p2 <- ggplot(dat, aes(year, log(cod_spr_ratio), color = SubDiv)) +
  geom_point() +
  stat_smooth() +
  scale_color_brewer(palette = "Dark2") + 
  ggtitle("Sprat") +
  theme(aspect.ratio = 3/4)

p1/p2

# Scaled variables
p3 <- dat %>% filter(cod_her_ratio_sc > -10 & cod_her_ratio_sc < 10) %>% 
  ggplot(., aes(year, cod_her_ratio_sc, color = SubDiv, fill = SubDiv)) +
  geom_point() +
  stat_smooth(alpha = 0.2) +
  scale_color_brewer(palette = "Dark2") + 
  scale_fill_brewer(palette = "Dark2") +
  labs(x = "Year", y = "cod:herring ratio") +
  theme(aspect.ratio = 3/4) + 
  facet_wrap(~SubDiv, ncol = 2) +
  guides(color = FALSE, fill = FALSE) +
  NULL

# Sprat ratio
p4 <- dat %>% filter(cod_spr_ratio_sc > -10 & cod_spr_ratio_sc < 10) %>% 
  ggplot(., aes(year, cod_spr_ratio_sc, color = SubDiv, fill = SubDiv)) +
  geom_point() +
  stat_smooth(alpha = 0.2) +
  scale_color_brewer(palette = "Dark2") + 
  scale_fill_brewer(palette = "Dark2") +
  labs(x = "Year", y = "cod:sprat ratio") +
  theme(aspect.ratio = 3/4) + 
  facet_wrap(~SubDiv, ncol = 2) +
  guides(color = FALSE, fill = FALSE) +
  NULL

p3+p4

ggsave("figures/supp/pelagics_ratio_time_series.png", width = 6.5, height = 6.5, dpi = 600)


# All separate
# Cod
p5 <- ggplot(dat, aes(year, ave_cpue_cod, color = SubDiv)) +
  geom_point() +
  stat_smooth(se = FALSE) +
  scale_color_brewer(palette = "Dark2") +
  ggtitle("Cod") +
  theme(aspect.ratio = 3/4)

# Herring
p6 <- ggplot(dat, aes(year, ave_abun_her, color = SubDiv, fill = SubDiv)) +
  geom_point(size = 1.1) +
  stat_smooth(alpha = 0.4, size = 0.7) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") + 
  theme_classic(base_size = 10) +
  theme(aspect.ratio = 3/4) +
  facet_wrap(~SubDiv, ncol = 2) +
  guides(color = FALSE, fill = FALSE) +
  labs(x = "Year", y = "Herring abundance (averaged across ICES rectangles)") +
  NULL

# Sprat ratio
p7 <- ggplot(dat, aes(year, ave_abun_spr, color = SubDiv, fill = SubDiv)) +
  geom_point(size = 1.1) +
  stat_smooth(alpha = 0.4, size = 0.7) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") + 
  theme_classic(base_size = 10) +
  theme(aspect.ratio = 3/4) +
  facet_wrap(~SubDiv, ncol = 2) +
  guides(color = FALSE, fill = FALSE) + 
  labs(x = "Year", y = "Sprat abundance (averaged across ICES rectangles)") +
  NULL

p6 + p7 

ggsave("figures/supp/pelagics_time_series.png", width = 6.5, height = 6.5, dpi = 600)
```

Marginal effect of oxygen

```{r marginal effects oxygen, message=FALSE, cache=TRUE}
# Prepare prediction data frame
d2 <- d %>% drop_na(oxy)
nd_oxy <- data.frame(oxy = seq(min(d2$oxy), max(d2$oxy), length.out = 100))

nd_oxy <- nd_oxy %>%
  mutate(year = 2003L,
         depth_sc = 0,
         oxy_sc = (oxy - mean(oxy))/sd(oxy),
         temp_sc = 0)

# Predict 
p_margin_oxy <- predict(m, newdata = nd_oxy, se_fit = TRUE, re_form = NA)

ggplot(p_margin_oxy, aes(oxy, exp(est),
  ymin = exp(est) - 1.96 * exp(est_se), ymax = exp(est) + 1.96 * exp(est_se))) +
  geom_line() + 
  geom_ribbon(alpha = 0.4)
```

Marginal effect of depth

```{r marginal effects depth, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_dep <- data.frame(depth = seq(min(d2$depth), max(d2$depth), length.out = 100))

nd_dep <- nd_dep %>%
  mutate(year = 2003L,
         depth_sc = (depth - mean(depth))/sd(depth),
         oxy_sc = 0,
         temp_sc = 0)

# Predict 
p_margin_dep <- predict(m, newdata = nd_dep, se_fit = TRUE, re_form = NA)

ggplot(p_margin_dep, aes(depth, exp(est),
  ymin = exp(est) - 1.96 * exp(est_se), ymax = exp(est) + 1.96 * exp(est_se))) +
  geom_line() + 
  geom_ribbon(alpha = 0.4)
```

Marginal effect of temperature

```{r marginal effects temperature, message=FALSE, cache=TRUE}
# Prepare prediction data frame
nd_temp <- data.frame(temp = seq(min(d$temp), max(d$temp), length.out = 100))

nd_temp <- nd_temp %>%
  mutate(year = 2003L,
         depth_sc = 0,
         oxy_sc = 0,
         temp_sc = (temp - mean(temp))/sd(temp))

# Predict 
p_margin_temp <- predict(m, newdata = nd_temp, se_fit = TRUE, re_form = NA)

ggplot(p_margin_temp, aes(temp, exp(est),
  ymin = exp(est) - 1.96 * exp(est_se), ymax = exp(est) + 1.96 * exp(est_se))) +
  geom_line() + 
  geom_ribbon(alpha = 0.4)
```
