---
title: "Collate CPUE data"
author: "Max Lindmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  df_print: paged
pdf_document: default
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.asp = 0.618,
  fig.align ='center'
)
```

# Intro
In this script, I load exchange data from datras and calculate catch of cod and flounder in unit kg/km^2 (with TVL gear), by correcting for gear dimensions, sweeplength and trawl speed, following Orio et al 2017. I also add oxygen, temperature and depth covariates, which I use for modelling this biomass index. 

# Load libraries
```{r, message=FALSE}
rm(list = ls())

library(tidyverse); theme_set(theme_light(base_size = 12))
library(readxl)
library(tidylog)
library(RCurl)
library(viridis)
library(RColorBrewer)
library(patchwork)
library(janitor)
library(icesDatras)
library(mapdata)
library(patchwork)
library(rgdal)
library(raster)
library(sf)
library(rgeos)
library(chron)
library(lattice)
library(ncdf4)
library(sdmTMB) # remotes::install_github("pbs-assess/sdmTMB")
library(marmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(mapplots)
library(geosphere)

world <- ne_countries(scale = "medium", returnclass = "sf")
```

#### Read data
```{r, message=FALSE}
# Read HH data
# bits_hh <- getDATRAS(record = "HH", survey = "BITS", years = 1991:2020, quarters = 4)
# write.csv(bits_hh, "data/DATRAS_exchange/bits_hh.csv")
bits_hh <- read.csv("data/DATRAS_exchange/bits_hh.csv")

# Read HL data
# bits_hl <- getDATRAS(record = "HL", survey = "BITS", years = 1991:2020, quarters = 4)
# write.csv(bits_hl, "data/DATRAS_exchange/bits_hl.csv")
bits_hl <- read.csv("data/DATRAS_exchange/bits_hl.csv")

# Read CA data
# bits_ca <- getDATRAS(record = "CA", survey = "BITS", years = 1991:2020, quarters = 4)
# write.csv(bits_ca, "data/DATRAS_exchange/bits_ca.csv")
bits_ca <- read.csv("data/DATRAS_exchange/bits_ca.csv")

# Read gear standardization data 
#sweep <- read.csv("data/from_ale/sweep_9116.csv", sep = ";", dec = ",", fileEncoding = "latin1")
sweep <- read.csv("data/from_ale/sweep_9118_ml.csv", sep = ";", fileEncoding = "latin1")
```

# Standardize catch data
##### Standardize ships
```{r, message=FALSE}
# Before creating a a new ID, make sure that countries and ships names use the same format
sort(unique(sweep$Ship))
sort(unique(bits_hh$Ship))
sort(unique(bits_hl$Ship))

# Change back to the old Ship name standard...
# https://vocab.ices.dk/?ref=315
# https://vocab.ices.dk/?ref=315
# Assumptions:
# SOL is Solea on ICES links above, and SOL1 is the older one of the two SOLs (1 and 2)
# DAN is Dana
# sweep %>% filter(Ship == "DANS") %>% distinct(Year, Country)
# sweep %>% filter(Ship == "DAN2") %>% distinct(Year)
# bits_hh %>% filter(Ship == "67BC") %>% distinct(Year, Country)
# sweep %>% filter(Ship == "DAN2") %>% distinct(Year)
# bits_hh %>% filter(Ship == "26D4") %>% distinct(Year) # Strange that 26DF doesn't extend far back. Which ship did the Danes use? Ok, I have no Danish data that old.
# bits_hh %>% filter(Country == "DK") %>% distinct(Year)

bits_hh <- bits_hh %>%
  mutate(Ship2 = fct_recode(Ship,
                            "SOL" = "06S1", 
                            "SOL2" = "06SL",
                            "DAN2" = "26D4",
                            "HAF" = "26HF",
                            "HAF" = "26HI",
                            "HAF" = "67BC",
                            "BAL" = "67BC",
                            "ARG" = "77AR",
                            "77SE" = "77SE",
                            "AA36" = "AA36",
                            "KOOT" = "ESLF",
                            "KOH" = "ESTM",
                            "DAR" = "LTDA",
                            "ATLD" = "RUJB",
                            "ATL" = "RUNT"), 
         Ship2 = as.character(Ship2)) %>% 
  mutate(Ship3 = ifelse(Country == "LV" & Ship2 == "BAL", "BALL", Ship2))

bits_hl <- bits_hl %>%
  mutate(Ship2 = fct_recode(Ship,
                            "SOL" = "06S1", 
                            "SOL2" = "06SL",
                            "DAN2" = "26D4",
                            "HAF" = "26HF",
                            "HAF" = "26HI",
                            "HAF" = "67BC",
                            "BAL" = "67BC",
                            "ARG" = "77AR",
                            "77SE" = "77SE",
                            "AA36" = "AA36",
                            "KOOT" = "ESLF",
                            "KOH" = "ESTM",
                            "DAR" = "LTDA",
                            "ATLD" = "RUJB",
                            "ATL" = "RUNT"), 
         Ship2 = as.character(Ship2)) %>% 
  mutate(Ship3 = ifelse(Country == "LV" & Ship2 == "BAL", "BALL", Ship2))

bits_ca <- bits_ca %>%
  mutate(Ship2 = fct_recode(Ship,
                            "SOL" = "06S1", 
                            "SOL2" = "06SL",
                            "DAN2" = "26D4",
                            "HAF" = "26HF",
                            "HAF" = "26HI",
                            "HAF" = "67BC",
                            "BAL" = "67BC",
                            "ARG" = "77AR",
                            "77SE" = "77SE",
                            "AA36" = "AA36",
                            "KOOT" = "ESLF",
                            "KOH" = "ESTM",
                            "DAR" = "LTDA",
                            "ATLD" = "RUJB",
                            "ATL" = "RUNT"), 
         Ship2 = as.character(Ship2)) %>% 
  mutate(Ship3 = ifelse(Country == "LV" & Ship2 == "BAL", "BALL", Ship2))

# Ok, which ships are missing in the exchange data?
unique(bits_hh$Ship3)[!unique(bits_hh$Ship3) %in% unique(sweep$Ship)]
# Swedish Ships and unidentified ships are NOT in the Sweep data
unique(sweep$Ship3)[!unique(sweep$Ship3) %in% unique(bits_hh$Ship3)]
# But all Sweep Ships are in the exchange data
```

##### Standardize countries
```{r, message=FALSE}
# Now check which country codes are used
sort(unique(sweep$Country))
sort(unique(bits_hh$Country))

# https://www.nationsonline.org/oneworld/country_code_list.htm#E
bits_hh <- bits_hh %>%
  mutate(Country = fct_recode(Country,
                              "DEN" = "DK",
                              "EST" = "EE",
                              "GFR" = "DE",
                              "LAT" = "LV",
                              "LTU" = "LT",
                              "POL" = "PL",
                              "RUS" = "RU",
                              "SWE" = "SE"),
         Country = as.character(Country))

bits_hl <- bits_hl %>%
  mutate(Country = fct_recode(Country,
                              "DEN" = "DK",
                              "EST" = "EE",
                              "GFR" = "DE",
                              "LAT" = "LV",
                              "LTU" = "LT",
                              "POL" = "PL",
                              "RUS" = "RU",
                              "SWE" = "SE"),
         Country = as.character(Country))

bits_ca <- bits_ca %>%
  mutate(Country = fct_recode(Country,
                              "DEN" = "DK",
                              "EST" = "EE",
                              "GFR" = "DE",
                              "LAT" = "LV",
                              "LTU" = "LT",
                              "POL" = "PL",
                              "RUS" = "RU",
                              "SWE" = "SE"),
         Country = as.character(Country))

# Gear? Are they the same?
sort(unique(bits_hh$Gear))
sort(unique(bits_hl$Gear))
sort(unique(sweep$Gear))

# Which gears are NOT in the sweep data?
unique(bits_hl$Gear)[!unique(bits_hl$Gear) %in% unique(sweep$Gear)] 
```

#### Create a simple ID that works across all exchange data
```{r, message=FALSE}
# Create ID column
bits_ca <- bits_ca %>% 
  mutate(IDx = paste(Year, Quarter, Country, Ship, Gear, StNo, HaulNo, sep = "."))

bits_hl <- bits_hl %>% 
  mutate(IDx = paste(Year, Quarter, Country, Ship, Gear, StNo, HaulNo, sep = "."))

bits_hh <- bits_hh %>% 
  mutate(IDx = paste(Year, Quarter, Country, Ship, Gear, StNo, HaulNo, sep = "."))

# Works like a haul-id
# bits_hh %>% group_by(IDx) %>% mutate(n = n()) %>% ungroup() %>% distinct(n)
```

#### Create the same unique haul-ID in the cpue data that I have in the sweep-file
```{r, message=FALSE}
bits_hl <- bits_hl %>% 
  mutate(haul.id = paste(Year, Quarter, Country, Ship3, Gear, StNo, HaulNo, sep = ":")) 

bits_hh <- bits_hh %>% 
  mutate(haul.id = paste(Year, Quarter, Country, Ship3, Gear, StNo, HaulNo, sep = ":")) 
```

#### Clean DATRAS EXCHANGE data
```{r, message=FALSE}
# Select just valid, additional and no oxygen hauls
bits_hh <- bits_hh %>%
  #filter(!Country == "SWE") %>% # I'll deal with Sweden later...
  filter(HaulVal %in% c("A","N","V"))

# Add ICES rectangle
bits_hh$Rect <- mapplots::ices.rect2(lon = bits_hh$ShootLong, lat = bits_hh$ShootLat)

# Add ICES subdivisions
shape <- shapefile("data/ICES_StatRec_mapto_ICES_Areas/StatRec_map_Areas_Full_20170124.shp")

pts <- SpatialPoints(cbind(bits_hh$ShootLong, bits_hh$ShootLat), 
                     proj4string = CRS(proj4string(shape)))

bits_hh$SD <- over(pts, shape)$Area_27

# Rename subdivisions to the more common names and do some more filtering (by sub div and area)
sort(unique(bits_hh$SD))

bits_hh <- bits_hh %>% 
  mutate(SD = factor(SD),
         SD = fct_recode(SD,
                         "20" = "3.a.20",
                         "21" = "3.a.21",
                         "22" = "3.c.22",
                         "23" = "3.b.23",
                         "24" = "3.d.24",
                         "25" = "3.d.25",
                         "26" = "3.d.26",
                         "27" = "3.d.27",
                         "28" = "3.d.28.1",
                         "28" = "3.d.28.2",
                         "29" = "3.d.29",
                         "30" = "3.d.30"),
         SD = as.character(SD)) 

# Now add the fishing line information from the sweep file (we need that later
# to standardize based on gear geometry). We add in the in the HH data and then
# transfer it to the other exchange data files when left_joining.
# Check which Fishing lines I have in the sweep data:
fishing_line <- sweep %>% group_by(Gear) %>% distinct(Fishing.line)

bits_hh <- left_join(bits_hh, fishing_line)
# sweep %>% group_by(Gear) %>% distinct(Fishing.line)
# bits_hh %>% group_by(Gear) %>% distinct(Fishing.line)
bits_hh$Fishing.line <- as.numeric(bits_hh$Fishing.line)

# Which gears do now have fishing line?
bits_hh$Fishing.line[is.na(bits_hh$Fishing.line)] <- -9
bits_hh %>% filter(Fishing.line == -9) %>% distinct(Gear)

# FROM the index files (Orio)
# FOT has 83
# GOV has 160
# ESB ??
# GRT ??

# Add these values:
bits_hh <- bits_hh %>% mutate(Fishing.line = ifelse(Gear == "FOT", 83, Fishing.line))
bits_hh <- bits_hh %>% mutate(Fishing.line = ifelse(Gear == "GOV", 160, Fishing.line))

# Now select the hauls in the HH data when subsetting the HL data
bits_hl <- bits_hl %>%
  filter(haul.id %in% bits_hh$haul.id)

# Match columns from the HH data to the HL and CA data
sort(unique(bits_hh$SD))
sort(colnames(bits_hh))
bits_hh_merge <- bits_hh %>% 
                       dplyr::select(SD, Rect, HaulVal, StdSpecRecCode, BySpecRecCode, Fishing.line,
                                     DataType, HaulDur, GroundSpeed, haul.id, IDx, ShootLat, ShootLong)

bits_hl <- left_join(dplyr::select(bits_hl, -haul.id), bits_hh_merge, by = "IDx")
bits_ca <- left_join(bits_ca, bits_hh_merge, by = "IDx")

# Now filter the subdivisions I want from all data sets
bits_hh <- bits_hh %>% filter(SD %in% c(24, 25, 26, 27, 28))
bits_hl <- bits_hl %>% filter(SD %in% c(24, 25, 26, 27, 28))
bits_ca <- bits_ca %>% filter(SD %in% c(24, 25, 26, 27, 28))
```

#### Filter species
```{r, message=FALSE}
hlcod <- bits_hl %>%
  filter(SpecCode %in% c("126436", "164712")) %>% 
  mutate(Species = "Gadus morhua")

hlfle <- bits_hl %>%
  filter(SpecCode %in% c("127141", "172894")) %>% 
  mutate(Species = "Platichthys flesus")
```

#### Prepare to add 0 catches
```{r, message=FALSE}
# Find common columns in the HH and HL data (here already subset by species)
comcol <- intersect(names(hlcod), names(bits_hh))

# What is the proportion of zero-catch hauls?
# Here we don't have zero catches
hlcod %>%
  group_by(haul.id, Year) %>%
  summarise(CPUEun_haul = sum(HLNoAtLngt)) %>% 
  ungroup() %>% 
  mutate(zero_catch = ifelse(CPUEun_haul == 0, "Y", "N")) %>% 
  distinct(zero_catch)

# Cod: Add 0s and then remove lines with SpecVal = 0 (first NA because we don't have a match in the HH, then make them 0 later)
hlcod0 <- full_join(hlcod, bits_hh[, comcol], by = comcol)

# No zeroes yet
hlcod0 %>%
  group_by(haul.id, Year) %>%
  summarise(CPUEun_haul = sum(HLNoAtLngt)) %>% 
  ungroup() %>% 
  mutate(zero_catch = ifelse(CPUEun_haul == 0, "Y", "N")) %>% 
  distinct(zero_catch) 

hlcod0$SpecVal[is.na(hlcod0$SpecVal)] <- "zeroCatch"

hlcod0$SpecVal <- factor(hlcod0$SpecVal)

hlcod0 <-  hlcod0 %>% filter(!SpecVal == "0")

# Add species again after merge
hlcod0$Species<-"Gadus morhua"


# Flounder: Add 0s, remove them if StdSpecRecCode !=1 and then remove lines with SpecVal = 0
hlfle0 <- full_join(hlfle, bits_hh[, comcol], by = comcol)

hlfle0 <- hlfle0[!(is.na(hlfle0$Species) & hlfle0$StdSpecRecCode != 1),] 

hlfle0$SpecVal[is.na(hlfle0$SpecVal)] <- "zeroCatch"
hlfle0$SpecVal <- factor(hlfle0$SpecVal)

hlfle0 <-  hlfle0 %>% filter(!SpecVal == "0")

hlfle0$Species<-"Platichthys flesus"

# Check number of hauls per species
hlcod0 %>% distinct(haul.id) %>% nrow()
hlfle0 %>% distinct(haul.id) %>% nrow()
```

#### Create (unstandardized) CPUE for `SpecVal=1`. If `DataType=C` then `CPUEun=HLNoAtLngt`, if `DataType=R` then `CPUEun=HLNoAtLngt/(HaulDur/60)`, if `DataType=S` then `CPUEun=(HLNoAtLngt*SubFactor)/(HaulDur/60)`. If `SpecVal="zeroCatch"` then `CPUEun=0`, if `SpecVal=4` we need to decide (no length measurements, only total catch). Note that here we also add zero CPUE if `SpecVal=="zeroCatch"`.

Then I will sum for the same haul the CPUE of the same length classes if they were sampled with different subfactors or with different sexes.

```{r, message=FALSE}
# Cod
hlcod0 <- hlcod0 %>%
  mutate(CPUEun = ifelse(SpecVal == "1" & DataType == "C",
                         HLNoAtLngt,
                         
                         ifelse(SpecVal == "1" & DataType == "R",
                                HLNoAtLngt/(HaulDur/60),
                                
                                ifelse(SpecVal == "1" & DataType == "S",
                                       (HLNoAtLngt*SubFactor)/(HaulDur/60),
                                       
                                       ifelse(SpecVal == "zeroCatch", 0, NA)))))

# Plot and fill by zero catch
hlcod0 %>%
  group_by(haul.id, Year) %>%
  summarise(CPUEun_haul = sum(CPUEun)) %>% 
  ungroup() %>% 
  mutate(zero_catch = ifelse(CPUEun_haul == 0, "Y", "N")) %>%
  group_by(Year, zero_catch) %>% 
  summarise(n = n()) %>% 
  ggplot(., aes(x = Year, y = n, fill = zero_catch)) +
  geom_bar(stat = "identity")

# Some rows have multiple rows per combination of length class and haul id,
# so we need to sum it up 
hlcod0 %>% group_by(LngtClass, haul.id) %>% mutate(n = n()) %>% ungroup() %>% distinct(n)
hlcod0 %>% group_by(LngtClass, haul.id) %>% mutate(n = n()) %>% ungroup() %>% filter(n == 2) %>% as.data.frame() %>% head(20)
test <- hlcod0 %>% group_by(LngtClass, haul.id) %>% mutate(n = n()) %>% ungroup() %>% filter(n == 2)
test_id <- test$haul.id[2]

hlcodL <- hlcod0 %>% 
  group_by(LngtClass, haul.id) %>% 
  mutate(CPUEun = sum(CPUEun)) %>%
  ungroup() %>% 
  mutate(id3 = paste(haul.id, LngtClass)) %>% 
  distinct(id3, .keep_all = TRUE) %>% 
  dplyr::select(-X, -id3) # Clean up a bit

# Check with an ID
# filter(hlcod0, haul.id == test_id)
# filter(hlcodL, haul.id == test_id) %>% as.data.frame()

# Do we still have 0 catches?
hlcodL %>%
  group_by(haul.id, Year) %>%
  summarise(CPUEun_haul = sum(CPUEun)) %>% 
  ungroup() %>% 
  mutate(zero_catch = ifelse(CPUEun_haul == 0, "Y", "N")) %>%
  group_by(Year, zero_catch) %>% 
  summarise(n = n()) %>% 
  ggplot(., aes(x = Year, y = n, fill = zero_catch)) +
  geom_bar(stat = "identity")

# Flounder
hlfle0 <- hlfle0 %>%
  mutate(CPUEun = ifelse(SpecVal == "1" & DataType == "C",
                         HLNoAtLngt,
                         
                         ifelse(SpecVal == "1" & DataType == "R",
                                HLNoAtLngt/(HaulDur/60),
                                
                                ifelse(SpecVal == "1" & DataType == "S",
                                       (HLNoAtLngt*SubFactor)/(HaulDur/60),
                                       
                                       ifelse(SpecVal == "zeroCatch", 0, NA)))))

# Sum up the CPUES if multiple per length class and haul
hlfleL <- hlfle0 %>% 
  group_by(LngtClass, haul.id) %>% 
  mutate(CPUEun = sum(CPUEun)) %>%
  ungroup() %>% 
  mutate(id3 = paste(haul.id, LngtClass)) %>% 
  distinct(id3, .keep_all = TRUE) %>% 
  dplyr::select(-X, -id3)
```

#### Get and add annual weight-length relationships from the CA data for both cod and flounder so that I can calculate CPUE in biomass rather than numbers further down
```{r, message=FALSE}
# Cod
bits_ca_cod <- bits_ca %>% 
  filter(SpecCode %in% c("164712", "126436") & Year < 2020) %>% 
  mutate(StNo = as.numeric(StNo)) %>% 
  mutate(Species = "Cod") %>% 
  mutate(ID = paste(Year, Quarter, Country, Ship, Gear, StNo, HaulNo, sep = "."))

# Now I need to copy rows with NoAtLngt > 1 so that 1 row = 1 ind
# First make a small test
# nrow(bits_ca_cod)
# test_id <- head(filter(bits_ca_cod, CANoAtLngt == 5))$ID[1]
# filter(bits_ca_cod, ID == test_id & CANoAtLngt == 5)

bits_ca_cod <- bits_ca_cod %>% map_df(., rep, .$CANoAtLngt)

# head(data.frame(filter(bits_ca_cod, ID == test_id & CANoAtLngt == 5)), 20)
# nrow(bits_ca_cod)
# Looks ok!

# Standardize length and drop NA weights (need that for condition)
bits_ca_cod <- bits_ca_cod %>% 
  drop_na(IndWgt) %>% 
  drop_na(LngtClass) %>% 
  filter(IndWgt > 0 & LngtClass > 0) %>%  # Filter positive length and weight
  mutate(weight_kg = IndWgt/1000) %>% 
  mutate(length_cm = ifelse(LngtCode == ".", 
                            LngtClass/10,
                            LngtClass)) # Standardize length ((https://vocab.ices.dk/?ref=18))

# Plot
ggplot(bits_ca_cod, aes(IndWgt, length_cm)) +
  geom_point() + 
  facet_wrap(~Year)

# Now extract the coefficients for each year (not bothering with outliers at the moment)
cod_intercept <- bits_ca_cod %>%
  split(.$Year) %>%
  purrr::map(~lm(log(IndWgt) ~ log(length_cm), data = .x)) %>%
  purrr::map_df(broom::tidy, .id = 'Year') %>%
  filter(term == "(Intercept)") %>% 
  mutate(a = exp(estimate)) %>% 
  mutate(Year = as.integer(Year)) %>% 
  dplyr::select(Year, a)

cod_slope <- bits_ca_cod %>%
  split(.$Year) %>%
  purrr::map(~lm(log(IndWgt) ~ log(length_cm), data = .x)) %>%
  purrr::map_df(broom::tidy, .id = 'Year') %>%
  filter(term == "log(length_cm)") %>% 
  mutate(Year = as.integer(Year)) %>% 
  rename("b" = "estimate") %>% 
  dplyr::select(Year, b)

# Flounder
bits_ca_fle <- bits_ca %>% 
  filter(SpecCode %in% c("127141", "172894") & Year < 2020) %>% 
  mutate(StNo = as.numeric(StNo)) %>% 
  mutate(Species = "Flounder") %>% 
  mutate(ID = paste(Year, Quarter, Country, Ship, Gear, StNo, HaulNo, sep = "."))

bits_ca_fle <- bits_ca_fle %>% map_df(., rep, .$CANoAtLngt)

# Standardize length and drop NA weights (need that for condition)
bits_ca_fle <- bits_ca_fle %>% 
  drop_na(IndWgt) %>% 
  drop_na(LngtClass) %>% 
  filter(IndWgt > 0 & LngtClass > 0) %>%  # Filter positive length and weight
  mutate(weight_kg = IndWgt/1000) %>% 
  mutate(length_cm = ifelse(LngtCode == ".", 
                            LngtClass/10,
                            LngtClass)) %>% # Standardize length ((https://vocab.ices.dk/?ref=18))
  mutate(keep = ifelse(LngtCode == "." & Year == 2008, "N", "Y")) %>%
  filter(keep == "Y") %>% 
  filter(length_cm < 70)

# Plot
ggplot(bits_ca_fle, aes(IndWgt, length_cm, color = LngtCode)) +
  geom_point() + 
  facet_wrap(~Year)

# Now extract the coefficients for each year (not bothering with outliers at the moment)
fle_intercept <- bits_ca_fle %>%
  split(.$Year) %>%
  purrr::map(~lm(log(IndWgt) ~ log(length_cm), data = .x)) %>%
  purrr::map_df(broom::tidy, .id = 'Year') %>%
  filter(term == "(Intercept)") %>% 
  mutate(a = exp(estimate)) %>% 
  mutate(Year = as.integer(Year)) %>% 
  dplyr::select(Year, a)

fle_slope <- bits_ca_fle %>%
  split(.$Year) %>%
  purrr::map(~lm(log(IndWgt) ~ log(length_cm), data = .x)) %>%
  purrr::map_df(broom::tidy, .id = 'Year') %>%
  filter(term == "log(length_cm)") %>% 
  mutate(Year = as.integer(Year)) %>% 
  rename("b" = "estimate") %>% 
  dplyr::select(Year, b)
```

#### Join the annual L-W relationships to the respective catch data to calculate CPUE in biomass not abundance
```{r, message=FALSE}
# These are the haul-data
# hlcodL
# hlfleL
hlcodL <- left_join(hlcodL, cod_intercept, by = "Year")
hlcodL <- left_join(hlcodL, cod_slope, by = "Year")

hlfleL <- left_join(hlfleL, fle_intercept, by = "Year")
hlfleL <- left_join(hlfleL, fle_slope, by = "Year")
```

#### Convert from CPUE in numbers to kg
```{r, message=FALSE}
# First standardize length to cm and then check how zero-catches are implemented at this stage
hlcodL <- hlcodL %>% 
  mutate(length_cm = ifelse(LngtCode == ".", 
                            LngtClass/10,
                            LngtClass)) # Standardize length ((https://vocab.ices.dk/?ref=18))
  
filter(hlcodL, length_cm == 0) # No such thing

# Now check if all rows where length is NA are the ones with zero catch!
hlcodL %>% 
  mutate(length2 = replace_na(length_cm, -9),
         no_length = ifelse(length2 < 0, "T", "F")) %>% 
  ggplot(., aes(length2, CPUEun, color = no_length)) + geom_point(alpha = 0.2) + facet_wrap(~no_length)

# Right, so all hauls with zero catch have NA length_cm. I don't have any NA catches
t <- hlcodL %>% drop_na(CPUEun)
t <- hlcodL %>% filter(CPUEun == 0)
t <- hlcodL %>% drop_na(length_cm)

# In other words, a zero catch is when the catch is zero and length_cm is NA
# In order to not get any NA CPUEs in unit biomass because length is NA (I want them instead
# to be 0, as the numbers-CPUE is), I will replace length_cm == NA with length_cm == 0 before
# calculating biomass cpue
hlcodL <- hlcodL %>% mutate(length_cm2 = replace_na(length_cm, 0))

# Standardize length in the haul-data and calculate weight
hlcodL <- hlcodL %>% 
  mutate(weight_kg = (a*length_cm2^b)/1000) %>% 
  mutate(CPUEun_kg = weight_kg*CPUEun)

# Plot and check it's correct also in this data
ggplot(hlcodL, aes(weight_kg, length_cm2)) +
  geom_point() + 
  facet_wrap(~Year)


# Now do the same for flounder
# First standardize length to cm and then check how zero-catches are implemented at this stage
hlfleL <- hlfleL %>% 
  mutate(length_cm = ifelse(LngtCode %in% c(".", "0"), 
                            LngtClass/10,
                            LngtClass)) # Standardize length (https://vocab.ices.dk/?ref=18)

filter(hlfleL, length_cm == 0) # No such thing

bits_ca_fle <- bits_ca_fle %>% 
  drop_na(IndWgt) %>% 
  drop_na(LngtClass) %>% 
  filter(IndWgt > 0 & LngtClass > 0) %>%  # Filter positive length and weight
  mutate(weight_kg = IndWgt/1000) %>% 
  mutate(length_cm = ifelse(LngtCode == ".", 
                            LngtClass/10,
                            LngtClass)) %>% # Standardize length ((https://vocab.ices.dk/?ref=18))
  mutate(keep = ifelse(LngtCode == "." & Year == 2008, "N", "Y")) %>%
  filter(keep == "Y") %>% 
  filter(length_cm < 70)

# Now check if all rows where length is NA are the ones with zero catch!
hlfleL %>% 
  mutate(length2 = replace_na(length_cm, -9),
         no_length = ifelse(length2 < 0, "T", "F")) %>% 
  ggplot(., aes(length2, CPUEun, color = no_length)) + geom_point(alpha = 0.2) + facet_wrap(~no_length)

hlfleL %>% mutate(length2 = replace_na(length_cm, -9)) %>% group_by(length2) %>% distinct(CPUEun) %>% arrange(CPUEun)

# Right, so all hauls with zero catch have NA length_cm. I don't have any NA catches
t <- hlfleL %>% drop_na(CPUEun)
# Well, 11 rows. I will remove them
hlfleL <- hlfleL %>% drop_na(CPUEun)
t <- hlfleL %>% filter(CPUEun == 0)
t <- hlfleL %>% drop_na(length_cm)

# In other words, a zero catch is when the catch is zero and length_cm is NA
# In order to not get any NA CPUEs in unit biomass because length is NA (I want them instead
# to be 0, as the numbers-CPUE is), I will replace length_cm == NA with length_cm == 0 before
# calculating biomass cpue
hlfleL <- hlfleL %>% mutate(length_cm2 = replace_na(length_cm, 0))

# Standardize length in the haul-data and calculate weight
hlfleL <- hlfleL %>% 
  mutate(weight_kg = (a*length_cm2^b)/1000) %>% 
  mutate(CPUEun_kg = weight_kg*CPUEun)

# Plot and check it's correct also in this data
ggplot(hlfleL, aes(weight_kg, length_cm2)) +
  geom_point() + 
  facet_wrap(~Year)

# Check
t <- hlfleL %>% drop_na(CPUEun_kg) # Should not have any NA in biomass-catch
t <- hlfleL %>% filter(CPUEun_kg == 0) # Should result in a few percent of rows (note this is not proportion of hauls, but rows)
t <- hlfleL %>% drop_na(length_cm2) # Should be no NA

# What is the proportion of zero-catch hauls?
cod_0plot <- hlcodL %>%
  group_by(haul.id, Year) %>%
  summarise(CPUEun_haul = sum(CPUEun)) %>% 
  ungroup() %>% 
  mutate(zero_catch = ifelse(CPUEun_haul == 0, "Y", "N")) %>% 
  group_by(Year, zero_catch) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  pivot_wider(names_from = zero_catch, values_from = n) %>% 
  mutate(prop_zero_catch_hauls = Y/(N+Y)) %>% 
  ggplot(., aes(Year, prop_zero_catch_hauls)) + geom_bar(stat = "identity") + 
  coord_cartesian(expand = 0, ylim = c(0, 1)) + 
  ggtitle("Cod")

# How many zero-catch hauls?
fle_0plot <- hlfleL %>%
  group_by(haul.id, Year) %>%
  summarise(CPUEun_haul = sum(CPUEun)) %>% 
  ungroup() %>% 
  mutate(zero_catch = ifelse(CPUEun_haul == 0, "Y", "N")) %>% 
  group_by(Year, zero_catch) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  pivot_wider(names_from = zero_catch, values_from = n) %>% 
  mutate(prop_zero_catch_hauls = Y/(N+Y)) %>% 
  ggplot(., aes(Year, prop_zero_catch_hauls)) + geom_bar(stat = "identity") + 
  coord_cartesian(expand = 0, ylim = c(0, 1)) + 
  ggtitle("Flounder")

cod_0plot / fle_0plot
```

#### Standardize according to Orio
To get unit: kg of fish caught by trawling for 1 h a standard bottom swept area of 0.45km2 using a TVL trawl with 75 m sweeps at the standard speed of three knots
```{r, message=FALSE}
# Remove hauls done with the TVL gear with a SweepLngt < 50 (these are calibration hauls, pers. com. Anders & Ale)
# And also hauls without length-information
# Remove pelagic gear
hlcodL <- hlcodL %>%
  mutate(SweepLngt2 = replace_na(SweepLngt, 50)) %>% 
  mutate(keep = ifelse(Gear == "TVL" & SweepLngt2 < 50, "N", "Y")) %>% 
  filter(keep == "Y") %>% 
  dplyr::select(-keep, -SweepLngt2) %>% 
  filter(!Gear == "PEL")
  
hlfleL <- hlfleL %>%
  mutate(SweepLngt2 = replace_na(SweepLngt, 50)) %>% 
  mutate(keep = ifelse(Gear == "TVL" & SweepLngt2 < 50, "N", "Y")) %>% 
  filter(keep == "Y") %>% 
  dplyr::select(-keep, -SweepLngt2) %>% 
  filter(!Gear == "PEL")

# Add in RS and RSA-values from the sweep file
# CPUE should be multiplied with RS and RSA to standardize to a relative speed and gear dimension.
# There is not a single file will all RS and RSA values. Instead they come in three files:
# - sweep (non-Swedish hauls between 1991-2016)
# - + calculated based on trawl speed and gear dimensions.
# I will join in the RS and RSA values from all sources, then standardize and filter
# away non-standardized hauls
# sort(unique(sweep$Year))
# sort(unique(sweep$Country))

# Since I don't have the sweep data for Swedish data, I have to calculate it from scratch using the 
# equation in Orio's spreadsheet

# First I will join in the sweep data, 
sweep_sel <- sweep %>% rename("haul.id" = "ï..haul.id") %>% dplyr::select(haul.id, RSA, RS)

hlcodL2 <- left_join(hlcodL, sweep_sel)
hlfleL2 <- left_join(hlfleL, sweep_sel)

hlcodL2 <- hlcodL2 %>%
  rename("RS_sweep" = "RS",
         "RSA_sweep" = "RSA") %>% 
  mutate(RS_sweep = as.numeric(RS_sweep),
         RSA_sweep = as.numeric(RSA_sweep))

hlfleL2 <- hlfleL2 %>%
  rename("RS_sweep" = "RS",
         "RSA_sweep" = "RSA") %>% 
  mutate(RS_sweep = as.numeric(RS_sweep),
         RSA_sweep = as.numeric(RSA_sweep))

sort(colnames(hlcodL2))
sort(colnames(hlfleL2))

# I will calculate a RS and RSA column in the catch data based on Ale's equation in the sweep file:
sort(unique(hlcodL2$GroundSpeed))
sort(unique(hlcodL2$Fishing.line))
sort(unique(hlcodL2$SweepLngt))

# First replace -9 in the columns I use for the calculations with NA so I don't end up with real numbers that are wrong!
hlcodL2 <- hlcodL2 %>% mutate(GroundSpeed = ifelse(GroundSpeed == -9, NA, GroundSpeed),
                              Fishing.line = ifelse(Fishing.line == -9, NA, Fishing.line),
                              SweepLngt = ifelse(SweepLngt == -9, NA, SweepLngt))

hlfleL2 <- hlfleL2 %>% mutate(GroundSpeed = ifelse(GroundSpeed == -9, NA, GroundSpeed),
                              Fishing.line = ifelse(Fishing.line == -9, NA, Fishing.line),
                              SweepLngt = ifelse(SweepLngt == -9, NA, SweepLngt))

# Now calculate correction factors
hlcodL2 <- hlcodL2 %>% mutate(RS_x = 3/GroundSpeed,
                              Horizontal.opening..m. = Fishing.line*0.67,
                              Swep.one.side..after.formula...meter = 0.258819045*SweepLngt, # SIN(RADIANS(15))
                              Size.final..m = Horizontal.opening..m. + (Swep.one.side..after.formula...meter*2),
                              Swept.area = (Size.final..m*3*1860)/1000000,
                              RSA_x = 0.45388309675081/Swept.area)

hlfleL2 <- hlfleL2 %>% mutate(RS_x = 3/GroundSpeed,
                              Horizontal.opening..m. = Fishing.line*0.67,
                              Swep.one.side..after.formula...meter = 0.258819045*SweepLngt, # SIN(RADIANS(15))
                              Size.final..m = Horizontal.opening..m. + (Swep.one.side..after.formula...meter*2),
                              Swept.area = (Size.final..m*3*1860)/1000000,
                              RSA_x = 0.45388309675081/Swept.area)

# Check EQ. is correct by recalculating it in the sweep file
sweep <- sweep %>% mutate(Horizontal.opening..m.2 = Fishing.line*0.67,
                          Swep.one.side..after.formula...meter2 = 0.258819045*SweepLngt, # SIN(RADIANS(15))
                          Size.final..m2 = Horizontal.opening..m.2 + (Swep.one.side..after.formula...meter2*2),
                          Swept.area2 = (Size.final..m2*3*1860)/1000000,
                          RSA_x = 0.45388309675081/Swept.area2)

sweep %>%
  drop_na() %>%
  ggplot(., aes(as.numeric(RSA), RSA_x)) + geom_point() + geom_abline(intercept = 0, slope = 1)

# Replace NAs with -1/3 (because ICES codes missing values as -9 and in the calculation above they get -1/3),
# so that I can filter them easily later
hlcodL2$RS_x[is.na(hlcodL2$RS_x)] <- -1/3
hlcodL2$RS_sweep[is.na(hlcodL2$RS_sweep)] <- -1/3
hlcodL2$RSA_x[is.na(hlcodL2$RSA_x)] <- -1/3
hlcodL2$RSA_sweep[is.na(hlcodL2$RSA_sweep)] <- -1/3

hlfleL2$RS_x[is.na(hlfleL2$RS_x)] <- -1/3
hlfleL2$RS_sweep[is.na(hlfleL2$RS_sweep)] <- -1/3
hlfleL2$RSA_x[is.na(hlfleL2$RSA_x)] <- -1/3
hlfleL2$RSA_sweep[is.na(hlfleL2$RSA_sweep)] <- -1/3

# Compare the difference correction factors (calculated vs imported from sweep file)
p1 <- ggplot(filter(hlcodL2, RS_x > 0), aes(RS_x)) + geom_histogram() + xlim(0.4, 1.76)
p2 <- ggplot(hlcodL2, aes(RSA_x)) + geom_histogram()
p3 <- ggplot(hlcodL2, aes(RS_sweep)) + geom_histogram() + xlim(0.4, 1.76)
p4 <- ggplot(hlcodL2, aes(RSA_sweep)) + geom_histogram()

(p1 + p2) / (p3 + p4)

p5 <- ggplot(filter(hlfleL2, RS_x > 0), aes(RS_x)) + geom_histogram() + xlim(0.4, 1.76)
p6 <- ggplot(hlfleL2, aes(RSA_x)) + geom_histogram()
p7 <- ggplot(hlfleL2, aes(RS_sweep)) + geom_histogram() + xlim(0.4, 1.76)
p8 <- ggplot(hlfleL2, aes(RSA_sweep)) + geom_histogram()

(p5 + p6) / (p7 + p8)

# Why do I have RSA values smaller than one? (eihter because sweep length is longer or gear is larger (GOV))
# Check if I can calculate the same RSA in sweep as that entered there.
# Ok, so the equation is correct. Which ID's have RSA < 1?
hlcodL2 %>% 
  filter(RSA_x < 1 & RSA_x > 0) %>%
  dplyr::select(Year, Country, Ship, Gear, haul.id, Horizontal.opening..m., Fishing.line,
                Swep.one.side..after.formula...meter, SweepLngt, Size.final..m, Swept.area, RSA_x) %>% 
  ggplot(., aes(RSA_x, fill = factor(SweepLngt))) + geom_histogram() + facet_wrap(~Gear, ncol = 1)

# Check if I have more than one unique RS or RSA value per haul, or if it's "either this or that"
# Filter positive in both columns
hlcodL2 %>% filter(RS_x > 0 & RS_sweep > 0) %>% ggplot(., aes(RS_x, RS_sweep)) +
  geom_point() + geom_abline(aes(intercept = 0, slope = 1), color = "red")

hlcodL2 %>% filter(RSA_x > 0 & RSA_sweep > 0) %>% ggplot(., aes(RSA_x, RSA_sweep)) +
  geom_point() + geom_abline(aes(intercept = 0, slope = 1), color = "red")

hlfleL2 %>% filter(RS_x > 0 & RS_sweep > 0) %>% ggplot(., aes(RS_x, RS_sweep)) +
  geom_point() + geom_abline(aes(intercept = 0, slope = 1), color = "red")

hlfleL2 %>% filter(RSA_x > 0 & RSA_sweep > 0) %>% ggplot(., aes(RSA_x, RSA_sweep)) +
  geom_point() + geom_abline(aes(intercept = 0, slope = 1), color = "red")

# Ok, there's on odd RS_x that is larger than 3. It didn't catch anything and speed is 0.8! Will remove
hlcodL2 <- hlcodL2 %>% filter(RS_x < 3)
hlfleL2 <- hlfleL2 %>% filter(RS_x < 3)

# Plot again
hlcodL2 %>% filter(RS_x > 0 & RS_sweep > 0) %>% ggplot(., aes(RS_x, RS_sweep)) +
  geom_point() + geom_abline(aes(intercept = 0, slope = 1), color = "red")

hlfleL2 %>% filter(RS_x > 0 & RS_sweep > 0) %>% ggplot(., aes(RS_x, RS_sweep)) +
  geom_point() + geom_abline(aes(intercept = 0, slope = 1), color = "red")

# They are the same when they overlap, so it does not matter which data source I use for RS and RSA values
# Make a single RS and RSA column

# Cod 
hlcodL3 <- hlcodL2 %>%
  mutate(RS = -99,
         RS = ifelse(RS_sweep > 0, RS_sweep, RS),
         RS = ifelse(RS < 0 & RS_x > 0, RS_x, RS)) %>% # Note that there are no NA i RS_x. This replaces all non-RS_sweep values -0.3, so I can filter positive later
  mutate(RSA = -99,
         RSA = ifelse(RSA_sweep > 0, RSA_sweep, RSA),
         RSA = ifelse(RSA < 0 & RSA_x > 0, RSA_x, RSA)) %>%
  filter(RS > 0) %>%
  filter(RSA > 0) %>% 
  mutate(RSRSA = RS*RSA)

# Plot
ggplot(hlcodL3, aes(RSRSA)) + geom_histogram()

hlfleL2 %>% filter(Country == "LAT") %>% distinct(Year) %>% arrange(Year)

# Flounder 
hlfleL3 <- hlfleL2 %>%
  mutate(RS = -999,
         RS = ifelse(RS_sweep > 0, RS_sweep, RS),
         RS = ifelse(RS < 0, RS_x, RS)) %>% # Note that there are no NA i RS_x. This replaces all non-RS_sweep values -0.3, so I can filter positive later
  mutate(RSA = -999,
         RSA = ifelse(RSA_sweep > 0, RSA_sweep, RSA),
         RSA = ifelse(RSA < 0, RSA_x, RSA)) %>% 
  filter(RS > 0) %>%
  filter(RSA > 0) %>% 
  mutate(RSRSA = RS*RSA)

# Test how many years of LAT data I miss out on because I can't standardize it.
# hlfleL2 %>%
#   mutate(RS = -999,
#          RS = ifelse(RS_sweep > 0, RS_sweep, RS),
#          RS = ifelse(RS < 0, RS_x, RS)) %>% # Note that there are no NA i RS_x. This replaces all non-RS_sweep values -0.3, so I can filter positive later
#   filter(RS > 0) %>% 
#   filter(Country == "LAT") %>% 
#   distinct(Year) %>% 
#   arrange(Year)
#   
# hlfleL2 %>%
#   mutate(RSA = -999,
#          RSA = ifelse(RSA_sweep > 0, RSA_sweep, RSA),
#          RSA = ifelse(RSA < 0, RSA_x, RSA)) %>% 
#   filter(RSA > 0) %>% 
#   filter(Country == "LAT") %>% 
#   distinct(Year) %>% 
#   arrange(Year)

# Plot
ggplot(hlcodL3, aes(RSRSA)) + geom_histogram()

# Standardize!
hlcodL3 <- hlcodL3 %>%
  mutate(CPUEst_kg = CPUEun_kg*RS*RSA,
         CPUEst = CPUEun*RS*RSA)

hlfleL3 <- hlfleL3 %>%
  mutate(CPUEst_kg = CPUEun_kg*RS*RSA,
         CPUEst = CPUEun*RS*RSA)
  
unique(is.na(hlcodL3$CPUEst_kg))
unique(is.na(hlcodL3$CPUEst))
min(hlcodL3$CPUEst_kg)
min(hlcodL3$CPUEst)

unique(is.na(hlfleL3$CPUEst_kg)) # Remove the few NA's here
hlfleL3 <- hlfleL3 %>% drop_na(CPUEst_kg)
unique(is.na(hlfleL3$CPUEst))
min(hlfleL3$CPUEst_kg) 
min(hlfleL3$CPUEst)

# Now calculate total CPUE per haul in Orios unit, then create the new unit, i.e.:
# convert from kg of fish caught by trawling for 1 h a standard bottom swept area of
# 0.45km2 (using a TVL trawl with 75 m sweeps at the standard speed of three knots) to....
# kg of fish per km^2 by dividing with 0.45

hlcodhaul <- hlcodL3 %>%
  group_by(haul.id) %>% 
  mutate(haul_cpue_kg = sum(CPUEst_kg),
         haul_cpue = sum(CPUEst),
         haul_cpue_kg_un = sum(CPUEun_kg),
         haul_cpue_un = sum(CPUEun),
         density = haul_cpue_kg/0.45,
         density_ab = haul_cpue/0.45) %>% 
  ungroup() %>% 
  distinct(haul.id, .keep_all = TRUE)

# t <- hlcodhaul %>% filter(haul_cpue_un == 0)
# t <- hlcodhaul %>% filter(!Country == "SWE") %>% filter(haul_cpue_un > 0)

hlflehaul <- hlfleL3 %>%
  group_by(haul.id) %>% 
  mutate(haul_cpue_kg = sum(CPUEst_kg),
         haul_cpue = sum(CPUEst),
         haul_cpue_kg_un = sum(CPUEun_kg),
         haul_cpue_un = sum(CPUEun),
         density = haul_cpue_kg/0.45,
         density_ab = haul_cpue/0.45) %>% 
  ungroup() %>% 
  distinct(haul.id, .keep_all = TRUE)

# Rename things and select specific columns
dat <- hlcodhaul %>% rename("year" = "Year",
                            "lat" = "ShootLat",
                            "lon" = "ShootLong",
                            "quarter" = "Quarter",
                            "ices_rect" = "Rect") %>% 
  dplyr::select(density, year, lat, lon, quarter, haul.id, IDx, ices_rect, SD)

# Now add in the flounder data in the cod data based on haul ID
hlflehaul_select <- hlflehaul %>% mutate(density_fle = density) %>% dplyr::select(density_fle, haul.id)

dat <- left_join(dat, hlflehaul_select, by = "haul.id") %>% drop_na(density_fle)

dat %>% arrange(desc(density_fle)) %>% data.frame() %>% head(50)

# Filter unrealistically large densities 
dat <- dat %>% filter(density_fle < 10000)

ggplot(dat, aes(density, density_fle)) + geom_point()
```

#### Compare cod data with Orio et al (2017)
```{r, message=FALSE}
# Read Orio data first for comparison:
test_cod <- hlcodhaul %>% filter(!Country == "SWE")
test_fle <- hlflehaul %>% filter(!Country == "SWE")

orio_cod <- read.csv("/Users/maxlindmark/Desktop/R_STUDIO_PROJECTS/ale_gear_standardization/datras_st_ale.csv")
orio_fle <- read.csv("/Users/maxlindmark/Desktop/R_STUDIO_PROJECTS/ale_gear_standardization/datras_fle_st_ale.csv")

orio_cod <- orio_cod %>%
  group_by(IDX) %>%
  mutate(haul_cpue_kg = sum(CPUEstBIOyear),
         haul_cpue = sum(CPUEst),
         haul_cpue_kg_un = sum(CPUEunBIOyear),
         haul_cpue_un = sum(CPUEun)) %>%
  ungroup() %>%
  distinct(IDX, .keep_all = TRUE)

orio_fle <- orio_fle %>%
  group_by(IDX) %>%
  mutate(haul_cpue_kg = sum(CPUEstBIOyear),
         haul_cpue = sum(CPUEst),
         haul_cpue_kg_un = sum(CPUEunBIOyear),
         haul_cpue_un = sum(CPUEun)) %>%
  ungroup() %>%
  distinct(IDX, .keep_all = TRUE)

# hlcodhaul %>% group_by(IDx) %>% mutate(n = n()) %>% ungroup() %>% distinct(n)

# Standardize data for easier plotting
test_cod <- test_cod %>%
  dplyr::select(haul_cpue_kg, haul_cpue, haul_cpue_kg_un, haul_cpue_un, Year, Ship3, Country, Gear) %>% 
  mutate(source = "Max",
         species = "Cod") %>% 
  rename("Ship" = "Ship3") %>% 
  filter(Year > 1992 & Year < 2017)

test_fle <- test_fle %>%
  dplyr::select(haul_cpue_kg, haul_cpue, haul_cpue_kg_un, haul_cpue_un, Year, Ship3, Country, Gear) %>% 
  mutate(source = "Max",
         species = "Fle") %>% 
  rename("Ship" = "Ship3") %>% 
  filter(Year > 1992 & Year < 2017)

orio_cod <- orio_cod %>%
  dplyr::select(haul_cpue_kg, haul_cpue, haul_cpue_kg_un, haul_cpue_un, year, vessel, IDX) %>% 
  mutate(source = "Ale",
         species = "Cod") %>% 
  rename("Year" = "year",
         "Ship" = "vessel") %>% 
  mutate(IDX2 = IDX,
         IDX3 = IDX) %>% 
  separate(IDX, sep = c(5:7), into = c("temp_year", "Quarter")) %>% 
  separate(temp_year, sep = c(4:5), into = c("Year", "scrap")) %>% 
  separate(IDX2, sep = 7, into = c("scrap2", "Country")) %>%
  separate(Country, sep = 3, into = c("Country", "scrap3")) %>%
  separate(IDX3, sep = 15, into = c("scrap4", "Gear")) %>%
  separate(Gear, sep = 3, into = c("Gear", "scrap5")) %>% 
  dplyr::select(-scrap, -scrap2, -scrap3, -scrap4, -scrap5) %>% 
  filter(Quarter == 4) %>% 
  mutate(Year = as.integer(Year)) %>% 
  mutate(haul_cpue_kg = haul_cpue_kg/1000,
         haul_cpue_kg_un = haul_cpue_kg_un/1000) %>% 
  filter(Year > 1992) %>% 
  filter(Country %in% unique(test_cod$Country))

orio_fle <- orio_fle %>% 
  dplyr::select(haul_cpue_kg, haul_cpue, haul_cpue_kg_un, haul_cpue_un, year, vessel, IDX) %>% 
  mutate(source = "Ale",
         species = "Fle") %>% 
  rename("Year" = "year",
         "Ship" = "vessel") %>% 
  mutate(IDX2 = IDX,
         IDX3 = IDX) %>% 
  separate(IDX, sep = c(5:7), into = c("temp_year", "Quarter")) %>% 
  separate(temp_year, sep = c(4:5), into = c("Year", "scrap")) %>% 
  separate(IDX2, sep = 7, into = c("scrap2", "Country")) %>%
  separate(Country, sep = 3, into = c("Country", "scrap3")) %>%
  separate(IDX3, sep = 15, into = c("scrap4", "Gear")) %>%
  separate(Gear, sep = 3, into = c("Gear", "scrap5")) %>% 
  dplyr::select(-scrap, -scrap2, -scrap3, -scrap4, -scrap5) %>% 
  filter(Quarter == 4) %>% 
  mutate(Year = as.integer(Year)) %>% 
  mutate(haul_cpue_kg = haul_cpue_kg/1000,
         haul_cpue_kg_un = haul_cpue_kg_un/1000) %>% 
  filter(Year > 1992) %>% 
  filter(Country %in% unique(test_cod$Country))

# sort(unique(test_cod$Country))
# sort(unique(orio_cod$Country))

# Check proportions of zero
t <- test_cod %>% filter(haul_cpue_kg > 0)
t <- orio_cod %>% filter(haul_cpue_kg > 0)

t <- test_fle %>% filter(haul_cpue_kg > 0)
t <- orio_fle %>% filter(haul_cpue_kg > 0)

test_full <- bind_rows(orio_fle, orio_cod, test_cod, test_fle)

# Check the non-standardized data for cod
# Raw abundance cpue
test_full %>% 
  filter(species == "Cod") %>% 
  ggplot(., aes(factor(Year), haul_cpue_un, color = source)) + 
  geom_point(size = 1, position = position_dodge(width = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90)) +
  NULL
  
test_full %>% 
  filter(species == "Cod") %>% 
  ggplot(., aes(haul_cpue_un, fill = source)) + 
  geom_histogram(position = position_dodge()) + 
  facet_wrap(~Year, scale = "free") +
  theme(axis.text.x = element_text(angle = 90)) + 
  NULL

# Mean abundance cpue
test_full %>% 
  filter(species == "Cod") %>% 
  group_by(Year, source) %>% 
  summarise(mean_cpue = mean(haul_cpue_un),
            sd_cpue = sd(haul_cpue_un)) %>% 
  ggplot(., aes(Year, mean_cpue, linetype = source, color = source)) + 
  geom_line(size = 1.1) +
  geom_point(size = 3) + 
  NULL

# Mean abundance cpue by country
test_full %>% 
  filter(species == "Cod") %>% 
  group_by(Year, source, Country) %>% 
  summarise(mean_cpue = mean(haul_cpue_un),
            sd_cpue = sd(haul_cpue_un)) %>% 
  ggplot(., aes(Year, mean_cpue, linetype = source, color = source)) + 
  geom_line(size = 1.1) +
  facet_wrap(~ Country) +
  NULL

# Mean abundance cpue for non-zero cathes
test_full %>% 
  filter(species == "Cod") %>% 
  filter(haul_cpue_un > 0) %>% 
  group_by(Year, source) %>% 
  summarise(mean_cpue = mean(haul_cpue_un),
            sd_cpue = sd(haul_cpue_un)) %>% 
  ggplot(., aes(Year, mean_cpue, linetype = source, color = source)) + 
  geom_line(size = 1.1) +
  geom_point(size = 3) + 
  NULL

# Now plot corrected biomass cpue
test_full %>% 
  filter(species == "Cod") %>% 
  ggplot(., aes(factor(Year), haul_cpue_kg, color = source)) + 
  geom_point(size = 1, position = position_dodge(width = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90)) +
  NULL

# Mean corrected biomass cpue
test_full %>% 
  filter(species == "Cod") %>% 
  group_by(Year, source) %>% 
  summarise(mean_cpue = mean(haul_cpue_kg),
            sd_cpue = sd(haul_cpue_kg)) %>% 
  ggplot(., aes(Year, mean_cpue, linetype = source, color = source)) + 
  geom_line(size = 1.1) +
  geom_point(size = 3) + 
  NULL


# Now check in on flounder
# Raw abundance cpue
test_full %>% 
  filter(species == "Fle") %>% 
  ggplot(., aes(factor(Year), haul_cpue_un, color = source)) + 
  geom_point(size = 1, position = position_dodge(width = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90)) +
  NULL
  
test_full %>% 
  filter(species == "Fle") %>% 
  ggplot(., aes(haul_cpue_un, fill = source)) + 
  geom_histogram(position = position_dodge()) + 
  facet_wrap(~Year, scale = "free") +
  theme(axis.text.x = element_text(angle = 90)) + 
  NULL

# Mean abundance cpue
test_full %>% 
  filter(species == "Fle") %>% 
  group_by(Year, source) %>% 
  summarise(mean_cpue = mean(haul_cpue_un),
            sd_cpue = sd(haul_cpue_un)) %>% 
  ggplot(., aes(Year, mean_cpue, linetype = source, color = source)) + 
  geom_line(size = 1.1) +
  geom_point(size = 3) + 
  NULL

# Ok, here we can see that Ale predicts an increase earlier, which we also saw 
# on the plot of raw catches as points
# Mean abundance cpue by country to see how this can arise
test_full %>% 
  filter(species == "Fle") %>% 
  group_by(Year, source, Country) %>% 
  summarise(mean_cpue = mean(haul_cpue_un),
            sd_cpue = sd(haul_cpue_un)) %>% 
  ggplot(., aes(Year, mean_cpue, linetype = source, color = source)) + 
  geom_line(size = 1.1) +
  facet_wrap(~ Country) +
  NULL

# Ok, after checking, the reason I don't have the older LAT data is because I don't
# have RS or RSA-values for those hauls, even though the catch data are in datras...
# Should probably check with Ale how he corrected those!

# Now plot corrected biomass cpue
test_full %>% 
  filter(species == "Fle") %>% 
  ggplot(., aes(factor(Year), haul_cpue_kg, color = source)) + 
  geom_point(size = 1, position = position_dodge(width = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90)) +
  NULL

# Mean corrected biomass cpue
test_full %>% 
  filter(species == "Fle") %>% 
  group_by(Year, source) %>% 
  summarise(mean_cpue = mean(haul_cpue_kg),
            sd_cpue = sd(haul_cpue_kg)) %>% 
  ggplot(., aes(Year, mean_cpue, linetype = source, color = source)) + 
  geom_line(size = 1.1) +
  geom_point(size = 3) + 
  NULL

# Then check lenght-weight relationships again
```

# Add in the environmental variables
#### Depth
```{r}
# Read the tifs
west <- raster("data/depth_geo_tif/D5_2018_rgb-1.tif")
#plot(west)

east <- raster("data/depth_geo_tif/D6_2018_rgb-1.tif")
# plot(east)

dep_rast <- raster::merge(west, east)

dat$depth <- extract(dep_rast, dat[, 4:3])

# Convert to depth (instead of elevation)
ggplot(dat, aes(depth)) + geom_histogram()
dat$depth <- (dat$depth - max(drop_na(dat)$depth)) *-1
ggplot(dat, aes(depth)) + geom_histogram()
```

#### Oxygen
```{r}
# Downloaded from here: https://resources.marine.copernicus.eu/?option=com_csw&view=details&product_id=BALTICSEA_REANALYSIS_BIO_003_012
# Extract raster points: https://gisday.wordpress.com/2014/03/24/extract-raster-values-from-points-using-r/comment-page-1/
# https://rpubs.com/boyerag/297592
# https://pjbartlein.github.io/REarthSysSci/netCDF.html#get-a-variable
# Open the netCDF file
ncin <- nc_open("data/NEMO_Nordic_SCOBI/dataset-reanalysis-scobi-monthlymeans_1610091357600.nc")

print(ncin)

# Get longitude and latitude
lon <- ncvar_get(ncin,"longitude")
nlon <- dim(lon)
head(lon)

lat <- ncvar_get(ncin,"latitude")
nlat <- dim(lat)
head(lat)

# Get time
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
nt
tunits

# Get oxygen
dname <- "o2b"

oxy_array <- ncvar_get(ncin,dname)
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(oxy_array)

# Get global attributes
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

# Convert time: split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])

# Here I deviate from the guide a little bit. Save this info:
dates <- chron(time, origin = c(tmonth, tday, tyear))

# Crop the date variable
months <- as.numeric(substr(dates, 2, 3))
years <- as.numeric(substr(dates, 8, 9))
years <- ifelse(years > 90, 1900 + years, 2000 + years)

# Replace netCDF fill values with NA's
oxy_array[oxy_array == fillvalue$value] <- NA

# We only use Quarter 4 in this analysis, so now we want to loop through each time step,
# and if it is a good month save it as a raster.
# First get the index of months that correspond to Q4
months

index_keep <- which(months > 9)

oxy_q4 <- oxy_array[, , index_keep]

months_keep <- months[index_keep]

years_keep <- years[index_keep]

# Now we have an array with only Q4 data...
# We need to now calculate the average within a year.
# Get a sequence that takes every third value between 1: number of months (length)
loop_seq <- seq(1, dim(oxy_q4)[3], by = 3)

# Create objects that will hold data
dlist <- list()
oxy_10 <- c()
oxy_11 <- c()
oxy_12 <- c()
oxy_ave <- c()

# Loop through the vector sequence with every third value, then take the average of
# three consecutive months (i.e. q4)
for(i in loop_seq) {
  
  oxy_10 <- oxy_q4[, , (i)]
  oxy_11 <- oxy_q4[, , (i + 1)]
  oxy_12 <- oxy_q4[, , (i + 2)]
  
  oxy_ave <- (oxy_10 + oxy_11 + oxy_12) / 3
  
  list_pos <- ((i/3) - (1/3)) + 1 # to get index 1:n(years)
  
  dlist[[list_pos]] <- oxy_ave
  
}

# Now name the lists with the year:
names(dlist) <- unique(years_keep)

# Now I need to make a loop where I extract the raster value for each year...
# The cpue data is called dat so far in this script

# Filter years in the cpue data frame to only have the years I have oxygen for
d_sub_oxy <- dat %>% filter(year %in% names(dlist)) %>% droplevels()

# Create data holding object
data_list <- list()

# ... And for the oxygen raster
raster_list <- list()

# Create factor year for indexing the list in the loop
d_sub_oxy$year_f <- as.factor(d_sub_oxy$year)

# Loop through each year and extract raster values for the cpue data points
for(i in unique(d_sub_oxy$year_f)) {
  
  # Set plot limits
  ymin = 54; ymax = 58; xmin = 12; xmax = 22

  # Subset a year
  oxy_slice <- dlist[[i]]
  
  # Create raster for that year (i)
  r <- raster(t(oxy_slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat),
              crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
  
  # Flip...
  r <- flip(r, direction = 'y')
  
  #plot(r, main = i)
  
  # Filter the same year (i) in the cpue data and select only coordinates
  d_slice <- d_sub_oxy %>% filter(year_f == i) %>% dplyr::select(lon, lat)
  
  # Make into a SpatialPoints object
  data_sp <- SpatialPoints(d_slice)
  
  # Extract raster value (oxygen)
  rasValue <- raster::extract(r, data_sp)
  
  # Now we want to plot the results of the raster extractions by plotting the cpue
  # data points over a raster and saving it for each year.
  # Make the SpatialPoints object into a raster again (for pl)
  df <- as.data.frame(data_sp)
  
  # Add in the raster value in the df holding the coordinates for the cpue data
  d_slice$oxy <- rasValue
  
  # Add in which year
  d_slice$year <- i
  
  # Create a index for the data last where we store all years (because our loop index
  # i is not continuous, we can't use it directly)
  index <- as.numeric(d_slice$year)[1] - 1992
  
  # Add each years' data in the list
  data_list[[index]] <- d_slice
  
  # Save to check each year is ok! First convert the raster to points for plotting
  # (so that we can use ggplot)
  map.p <- rasterToPoints(r)
  
  # Make the points a dataframe for ggplot
  df_rast <- data.frame(map.p)
  
  # Rename y-variable and add year
  df_rast <- df_rast %>% rename("oxy" = "layer") %>% mutate(year = i)
  
  # Add each years' raster data frame in the list
  raster_list[[index]] <- df_rast
  
  # Make appropriate column headings
  colnames(df_rast) <- c("Longitude", "Latitude", "oxy")
  
  # Now make the map
  ggplot(data = df_rast, aes(y = Latitude, x = Longitude)) +
    geom_raster(aes(fill = oxy)) +
    geom_point(data = d_slice, aes(x = lon, y = lat, fill = oxy),
               color = "black", size = 5, shape = 21) +
    theme_bw() +
    geom_sf(data = world, inherit.aes = F, size = 0.2) +
    coord_sf(xlim = c(xmin, xmax),
             ylim = c(ymin, ymax)) +
    scale_colour_gradientn(colours = rev(terrain.colors(10)),
                           limits = c(-200, 400)) +
    scale_fill_gradientn(colours = rev(terrain.colors(10)),
                         limits = c(-200, 400)) +
    NULL
  
  ggsave(paste("figures/supp/cpue_oxygen_rasters/", i,".png", sep = ""),
         width = 6.5, height = 6.5, dpi = 600)
  
}

# Now create a data frame from the list of all annual values
big_dat_oxy <- dplyr::bind_rows(data_list)
big_raster_dat_oxy <- dplyr::bind_rows(raster_list)

# Plot data, looks like there's big inter-annual variation but a negative trend over time
big_raster_dat_oxy %>%
  group_by(year) %>%
  drop_na(oxy) %>%
  summarise(mean_oxy = mean(oxy)) %>%
  mutate(year_num = as.numeric(year)) %>%
  ggplot(., aes(year_num, mean_oxy)) +
  geom_point(size = 2) +
  stat_smooth(method = "lm") +
  NULL

big_raster_dat_oxy %>%
  group_by(year) %>%
  drop_na(oxy) %>%
  mutate(dead = ifelse(oxy < 0, "Y", "N")) %>%
  filter(dead == "Y") %>%
  mutate(n = n(),
         year_num = as.numeric(year)) %>%
  ggplot(., aes(year_num, n)) +
  geom_point(size = 2) +
  stat_smooth(method = "lm") +
  NULL

# Now add in the new oxygen column in the original data:
str(d_sub_oxy)
str(big_dat_oxy)

# Create an ID for matching the oxygen data with the cpue data
dat$id_oxy <- paste(dat$year, dat$lon, dat$lat, sep = "_")
big_dat_oxy$id_oxy <- paste(big_dat_oxy$year, big_dat_oxy$lon, big_dat_oxy$lat, sep = "_")

# Which id's are not in the cpue data (dat)?
ids <- dat$id_oxy[!dat$id_oxy %in% c(big_dat_oxy$id_oxy)]

unique(ids)

# Select only the columns we want to merge
big_dat_sub_oxy <- big_dat_oxy %>% dplyr::select(id_oxy, oxy)

# Remove duplicate ID (one oxy value per id)
big_dat_sub_oxy2 <- big_dat_sub_oxy %>% distinct(id_oxy, .keep_all = TRUE)
# big_dat_sub_oxy %>% group_by(id_oxy) %>% mutate(n = n()) %>% arrange(desc(n))

# Join the data with raster-derived oxygen with the full cpue data
dat <- left_join(dat, big_dat_sub_oxy2, by = "id_oxy")

# Now the unit of oxygen is mmol/m3. I want it to be ml/L. The original model is in unit ml/L
# and it's been converted by the data host. Since it was converted without accounting for
# pressure or temperature, I can simply use the following conversion factor:
# 1 ml/l = 103/22.391 = 44.661 μmol/l -> 1 ml/l = 0.044661 mmol/l = 44.661 mmol/m^3 -> 0.0223909 ml/l = 1mmol/m^3
# https://ocean.ices.dk/tools/unitconversion.aspx

dat$oxy <- dat$oxy * 0.0223909

# Drop NA oxygen
dat <- dat %>% drop_na(oxy)
```

#### Temperature
```{r}
# Open the netCDF file
ncin <- nc_open("data/NEMO_Nordic_SCOBI/dataset-reanalysis-nemo-monthlymeans_1608127623694.nc")

print(ncin)

# Get longitude and latitude
lon <- ncvar_get(ncin,"longitude")
nlon <- dim(lon)
head(lon)

lat <- ncvar_get(ncin,"latitude")
nlat <- dim(lat)
head(lat)

# Get time
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
nt
tunits

# Get temperature
dname <- "bottomT"

temp_array <- ncvar_get(ncin,dname)
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(temp_array)

# Get global attributes
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

# Convert time: split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])

# Here I deviate from the guide a little bit. Save this info:
dates <- chron(time, origin = c(tmonth, tday, tyear))

# Crop the date variable
months <- as.numeric(substr(dates, 2, 3))
years <- as.numeric(substr(dates, 8, 9))
years <- ifelse(years > 90, 1900 + years, 2000 + years)

# Replace netCDF fill values with NA's
temp_array[temp_array == fillvalue$value] <- NA

# We only use Quarter 4 in this analysis, so now we want to loop through each time step,
# and if it is a good month save it as a raster.
# First get the index of months that correspond to Q4
months

index_keep <- which(months > 9)

# Quarter 4 by keeping months in index_keep
temp_q4 <- temp_array[, , index_keep]

months_keep <- months[index_keep]

years_keep <- years[index_keep]

# Now we have an array with only Q4 data...
# We need to now calculate the average within a year.
# Get a sequence that takes every third value between 1: number of months (length)
loop_seq <- seq(1, dim(temp_q4)[3], by = 3)

# Create objects that will hold data
dlist <- list()
temp_10 <- c()
temp_11 <- c()
temp_12 <- c()
temp_ave <- c()

# Loop through the vector sequence with every third value, then take the average of
# three consecutive months (i.e. q4)
for(i in loop_seq) {
  
  temp_10 <- temp_q4[, , (i)]
  temp_11 <- temp_q4[, , (i + 1)]
  temp_12 <- temp_q4[, , (i + 2)]
  
  temp_ave <- (temp_10 + temp_11 + temp_12) / 3
  
  list_pos <- ((i/3) - (1/3)) + 1 # to get index 1:n(years)
  
  dlist[[list_pos]] <- temp_ave
  
}

# Now name the lists with the year:
names(dlist) <- unique(years_keep)

# Now I need to make a loop where I extract the raster value for each year...
# The cpue data is called dat so far in this script

# Filter years in the cpue data frame to only have the years I have temperature for
d_sub_temp <- dat %>% filter(year %in% names(dlist)) %>% droplevels()

# Create data holding object
data_list <- list()

# ... And for the temperature raster
raster_list <- list()

# Create factor year for indexing the list in the loop
d_sub_temp$year_f <- as.factor(d_sub_temp$year)

# Loop through each year and extract raster values for the cpue data points
for(i in unique(d_sub_temp$year_f)) {
  
  # Subset a year
  temp_slice <- dlist[[i]]
  
  # Create raster for that year (i)
  r <- raster(t(temp_slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat),
              crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
  
  # Flip...
  r <- flip(r, direction = 'y')
  
  #plot(r, main = i)
  
  # Filter the same year (i) in the cpue data and select only coordinates
  d_slice <- d_sub_temp %>% filter(year_f == i) %>% dplyr::select(lon, lat)
  
  # Make into a SpatialPoints object
  data_sp <- SpatialPoints(d_slice)
  
  # Extract raster value (temperature)
  rasValue <- raster::extract(r, data_sp)
  
  # Now we want to plot the results of the raster extractions by plotting the cpue
  # data points over a raster and saving it for each year.
  # Make the SpatialPoints object into a raster again (for pl)
  df <- as.data.frame(data_sp)
  
  # Add in the raster value in the df holding the coordinates for the cpue data
  d_slice$temp <- rasValue
  
  # Add in which year
  d_slice$year <- i
  
  # Create a index for the data last where we store all years (because our loop index
  # i is not continuous, we can't use it directly)
  index <- as.numeric(d_slice$year)[1] - 1992
  
  # Add each years' data in the list
  data_list[[index]] <- d_slice
  
  # Save to check each year is ok! First convert the raster to points for plotting
  # (so that we can use ggplot)
  map.p <- rasterToPoints(r)
  
  # Make the points a dataframe for ggplot
  df_rast <- data.frame(map.p)
  
  # Rename y-variable and add year
  df_rast <- df_rast %>% rename("temp" = "layer") %>% mutate(year = i)
  
  # Add each years' raster data frame in the list
  raster_list[[index]] <- df_rast
  
  # Make appropriate column headings
  colnames(df_rast) <- c("Longitude", "Latitude", "temp")
  
  # Now make the map
  ggplot(data = df_rast, aes(y = Latitude, x = Longitude)) +
    geom_raster(aes(fill = temp)) +
    geom_point(data = d_slice, aes(x = lon, y = lat, fill = temp),
               color = "black", size = 5, shape = 21) +
    theme_bw() +
    geom_sf(data = world, inherit.aes = F, size = 0.2) +
    coord_sf(xlim = c(min(dat$lon), max(dat$lon)),
             ylim = c(min(dat$lat), max(dat$lat))) +
    scale_colour_gradientn(colours = rev(terrain.colors(10)),
                           limits = c(2, 17)) +
    scale_fill_gradientn(colours = rev(terrain.colors(10)),
                         limits = c(2, 17)) +
    NULL
  
  ggsave(paste("figures/supp/cpue_temp_rasters/", i,".png", sep = ""),
         width = 6.5, height = 6.5, dpi = 600)
  
}

# Now create a data frame from the list of all annual values
big_dat_temp <- dplyr::bind_rows(data_list)
big_raster_dat_temp <- dplyr::bind_rows(raster_list)

big_dat_temp %>% drop_na(temp) %>% summarise(max = max(temp))
big_dat_temp %>% drop_na(temp) %>% summarise(min = min(temp))

# Plot data, looks like there's big inter-annual variation but a positive trend
big_raster_dat_temp %>%
  group_by(year) %>%
  drop_na(temp) %>%
  summarise(mean_temp = mean(temp)) %>%
  mutate(year_num = as.numeric(year)) %>%
  ggplot(., aes(year_num, mean_temp)) +
  geom_point(size = 2) +
  stat_smooth(method = "lm") +
  NULL

# Now add in the new temperature column in the original data:
str(d_sub_temp)
str(big_dat_temp)

# Create an ID for matching the temperature data with the cpue data
dat$id_temp <- paste(dat$year, dat$lon, dat$lat, sep = "_")
big_dat_temp$id_temp <- paste(big_dat_temp$year, big_dat_temp$lon, big_dat_temp$lat, sep = "_")

# Which id's are not in the cpue data (dat)? (It's because I don't have those years, not about the location)
ids <- dat$id_temp[!dat$id_temp %in% c(big_dat_temp$id_temp)]

unique(ids)

# Select only the columns we want to merge
big_dat_sub_temp <- big_dat_temp %>% dplyr::select(id_temp, temp)

# Remove duplicate ID (one temp value per id)
big_dat_sub_temp2 <- big_dat_sub_temp %>% distinct(id_temp, .keep_all = TRUE)

# Join the data with raster-derived oxygen with the full cpue data
dat <- left_join(dat, big_dat_sub_temp2, by = "id_temp")

colnames(dat)

dat <- dat %>% dplyr::select(-id_temp, -id_oxy)

# Drop NA temp
dat <- dat %>% drop_na(temp)
```

# Add UTM coords
```{r}
# First add UTM coords
# Add UTM coords
# Function
LongLatToUTM <- function(x, y, zone){
  xy <- data.frame(ID = 1:length(x), X = x, Y = y)
  coordinates(xy) <- c("X", "Y")
  proj4string(xy) <- CRS("+proj=longlat +datum=WGS84")  ## for example
  res <- spTransform(xy, CRS(paste("+proj=utm +zone=",zone," ellps=WGS84",sep='')))
  return(as.data.frame(res))
}

utm_coords <- LongLatToUTM(dat$lon, dat$lat, zone = 33)
dat$X <- utm_coords$X/1000 # for computational reasons
dat$Y <- utm_coords$Y/1000 # for computational reasons
```

# Save data
```{r}
write.csv(dat, file = "data/for_analysis/mdat_cpue.csv", row.names = FALSE)
```
